---
title: 
csl: plos.csl
fontsize: 10pt
output:
  pdf_document:
    includes:
      in_header: latex/midyr-header.tex
    keep_tex: yes
  html_document:
    df_print: paged
link-citations: yes
geometry: left=2cm, right=2cm, top=2cm, bottom=3cm, footskip = .5cm
subparagraph: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(scholar)
library(RefManageR)
```

## Element 1: Research (35) 

### Objective: Support the development of ecosystem approaches to fishery management/Integrated Ecosystem Assessment (IEA) and Management Strategy Evaluation (MSE) for the U.S. Northeast Continental Shelf

1. Conduct research on approaches to for conducting and implementing ecosystem approaches and IEAs.

1. Contribute to the development and implementation of multispecies and ecosystem MSE operating models.

1. Communicate research results through peer-reviewed manuscripts, reports, and oral/poster presentations at scientific meetings

```{r}

if(!file.exists("gaichas.bib")){
  
scholar::get_publications("C_gOAZ4AAAAJ&hl") %>% 
  filter(year > 2018) %>% 
  arrange(desc(year)) %>%
  distinct(title, .keep_all = TRUE) %>%
  transmute(bibtype = "Article", author = as.character(author),
            title = as.character(title),
            journaltitle = as.character(journal), year, key = row_number()) %>% 
  RefManageR::as.BibEntry() %>% 
  RefManageR::WriteBib("gaichas.bib")
  
}

bibliography_entries("gaichas.bib")
```


1. Review scientific manuscripts, research proposals, and job completion reports as requested.



## Element 2: Program support (15) 

### Objective: Provide analytical support for EDAB and MSE research initiatives

1.  Provide analytical support for Ecosystem Approaches 

1.  Provide analytical support for Management Strategy Evaluation

1. Serve as mentor for Center staff, graduate students, and post-doctoral researchers as appropriate.


## ELement 3: Products and Services (30) 

### Objective: Provide accurate and timely information in support of Ecosystem Approaches to Management/Integrated Ecosystem Assessments and Management Strategy Evaluation

1. Ecosystem Approaches and IEA Products and Services 

1. Management Strategy Evaluation Products and Services

## Element 4: Planning and evaluation (20) for Ecosystem Approaches and MSE

1. Serve on committees, steering groups and task forces 

1. Actively promote interactions between the Ecosystem Assessment Program with other Center Divisions, the Center Directorate, and the interested public. Actively promote interdisciplinary MSE within and beyond the Center

1. Interact with stakeholders to disseminate knowledge via courses, seminars, lectures and workshops as appropriate


