---
address: NOAA NMFS Northeast Fisheries Science Center
date: "`r format(Sys.time(), '%B %Y')`"
email: Sarah.Gaichas@noaa.gov
github: sgaichas
name: Sarah
output:
  vitae::awesomecv:
    keep_tex: yes
  vitae::moderncv: default
phone: +1 508 495 2016
position: Research Fishery Biologist
surname: Gaichas
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
```

# Education

```{r}
library(tibble)

orcid_data <- do.call("rbind",
    rorcid::orcid_educations("0000-0002-5788-3073")$`0000-0002-5788-3073`$`affiliation-group`$summaries
)

orcid_data %>%
  detailed_entries(
    what = `education-summary.role-title`,
    when = glue::glue("{`education-summary.start-date.year.value`} - {`education-summary.end-date.year.value`}"),
    with = `education-summary.organization.name`,
    where = `education-summary.organization.address.city`
  )

# tribble(
#   ~ Degree, ~ Year, ~ Institution, ~ Where,
#   "Informal studies", "1889-91", "Flying University", "Warsaw, Poland",
#   "Master of Physics", "1893", "Sorbonne Université", "Paris, France",
#   "Master of Mathematics", "1894", "Sorbonne Université", "Paris, France"
# ) %>% 
#   detailed_entries(Degree, Year, Institution, Where)

```

# Graduate Supervision and Examination

**Adjunct professor**  *University of Massachusetts, Dartmouth* 

One current UMassD PhD. student committee.   
One current Ph.D. student committee at the College of William and Mary (Virginia Institute of Marine Science). 

2020  External examiner for a PhD student at University of Iceland  
2019  (Exam) PhD student committee at the University of Massachusetts, Dartmouth   
2019  External examiner for a PhD student at Dalhousie University  
2016	External examiner for a PhD student at the University of Cape Town   
2016	(Exam) PhD student committee at the University of Washington   
2015	(Exam) PhD student committee at Oregon State University  

# Experience

**Research Fishery Biologist**                   		*NOAA/NMFS Northeast Fisheries Sci. Center, Woods Hole, MA*

	Ecosystem Assessment, MSE		September 2011 - present
Contribute to integrated ecosystem assessments and management strategy evaluations (MSEs) for the Northeast U.S. continental shelf by analyzing research survey, fishery, oceanographic, and economic data, and by developing and testing multispecies assessment models and ecosystem indicators. Conduct research on ecosystem based management reference points balancing multiple objectives. Chair National and Northeast MSE working groups to build capacity and coordinate work plans. Provide results in peer-reviewed literature and to the New England and Mid-Atlantic Fishery Management Councils. 

**Research Fishery Biologist**                   		*NOAA/NMFS Alaska Fisheries Science Center, Seattle, WA*

	Ecosystem Modeling			June 2006 – September 2011
Analyzed food web, fishery, and research survey data and developed ecosystem models to evaluate the cumulative effects of fishing, climate change, and other factors on the marine ecosystem. 

	Stock Assessment			November 1998 – June 2006
Assessed the condition of commercially exploited fish stocks through analysis of fishery and research survey data combined with population dynamics modeling. 

	Observer Program			March 1997 – October 1999
Conducted statistical analyses of fishery observer data, evaluated present and proposed data collection methods through directed research, and examined uses of observer data in multiple applications.  

# Research, Science, and Management Organizations

* Mid-Atlantic FMC Scientific and Statistical Committee, (2014-present); determining Allowable Biological Catch (ABC) advice for fishery management and other scientific support. 
* Member and past co-chair of the ICES Working Group on North Atlantic Regional Seas; developing ecosystem indicators and integrated ecosystem assessments for Canadian and U.S. waters.
* Co-Chair of the ICES Working Group on Multispecies Assessment Methods: reviewing and contributing to multispecies and ecosystem modeling progress across ICES regions. 
* Advisory partner in the European Commission's PANDORA project (2018-2022), which is designing toolsets for new dynamic ocean resource assessments and exploitation.

# Publications

```{r bib, warning=FALSE, results='asis'}

if(!file.exists("gaichas.bib")){
scholar::get_publications("C_gOAZ4AAAAJ&hl") %>% 
  #filter(year > 2014) %>% 
  arrange(desc(year)) %>%
  distinct(title, .keep_all = TRUE) %>%
  transmute(bibtype = "Article", author = as.character(author),
            title = as.character(title),
            journaltitle = as.character(journal), year, key = row_number()) %>% 
  RefManageR::as.BibEntry() %>% 
  RefManageR::WriteBib("gaichas.bib")
}

# this bib from zotero; google scholar brought in conference papers, etc
bib <- RefManageR::ReadBib("gaichas.bib")
#sort(bib, sorting = "ydnt")
RefManageR::BibOptions(sorting = "ydnt", 
                       bib.style = "authoryear", 
                       max.names = 50,
                       style = "markdown",
                       no.print.fields = c("abstract", "copyright", "issn", "url", "language","urldate","month","file", "shorttitle", "note", "keywords"))
RefManageR::NoCite(bib)

newline <- list(rep("  \n", length(bib)))
RefManageR::BibEntry()
#bib <- mapply(append, RefManageR::BibEntry(bibtype = "doi"), newline, SIMPLIFY = FALSE)
RefManageR::PrintBibliography(bib)

#bibliography_entries("gaichas.bib")%>%
#  select(author, year, title, journal, pages, url) %>%
#  arrange(desc(year, author$family))
```

# Scientific Presentations 

Speaker, moderator, and symposium organizer at regional to international professional meetings (AFS, ASLO, Conservation Biology, ESA, GLOBEC, IMCC, World Fisheries Congress) 1995-present. Invited speaker for public lecture series, teacher workshops, and university courses and departmental seminars.

