%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode
% Awesome CV LaTeX Template for CV/Resume
%
% This template has been downloaded from:
% https://github.com/posquit0/Awesome-CV
%
% Author:
% Claud D. Park <posquit0.bj@gmail.com>
% http://www.posquit0.com
%
%
% Adapted to be an Rmarkdown template by Mitchell O'Hara-Wild
% 23 November 2018
%
% Template license:
% CC BY-SA 4.0 (https://creativecommons.org/licenses/by-sa/4.0/)
%
%-------------------------------------------------------------------------------
% CONFIGURATIONS
%-------------------------------------------------------------------------------
% A4 paper size by default, use 'letterpaper' for US letter
\documentclass[11pt, a4paper]{awesome-cv}

% Configure page margins with geometry
\geometry{left=1.4cm, top=.8cm, right=1.4cm, bottom=1.8cm, footskip=.5cm}

% Specify the location of the included fonts
\fontdir[fonts/]

% Color for highlights
% Awesome Colors: awesome-emerald, awesome-skyblue, awesome-red, awesome-pink, awesome-orange
%                 awesome-nephritis, awesome-concrete, awesome-darknight

\colorlet{awesome}{awesome-red}

% Colors for text
% Uncomment if you would like to specify your own color
% \definecolor{darktext}{HTML}{414141}
% \definecolor{text}{HTML}{333333}
% \definecolor{graytext}{HTML}{5D5D5D}
% \definecolor{lighttext}{HTML}{999999}

% Set false if you don't want to highlight section with awesome color
\setbool{acvSectionColorHighlight}{true}

% If you would like to change the social information separator from a pipe (|) to something else
\renewcommand{\acvHeaderSocialSep}{\quad\textbar\quad}

\def\endfirstpage{\newpage}

%-------------------------------------------------------------------------------
%	PERSONAL INFORMATION
%	Comment any of the lines below if they are not required
%-------------------------------------------------------------------------------
% Available options: circle|rectangle,edge/noedge,left/right

\name{Sarah}{Gaichas}

\position{Research Fishery Biologist}
\address{NOAA NMFS Northeast Fisheries Science Center}

\mobile{+1 508 495 2016}
\email{\href{mailto:Sarah.Gaichas@noaa.gov}{\nolinkurl{Sarah.Gaichas@noaa.gov}}}
\github{sgaichas}

% \gitlab{gitlab-id}
% \stackoverflow{SO-id}{SO-name}
% \skype{skype-id}
% \reddit{reddit-id}


\usepackage{booktabs}

\providecommand{\tightlist}{%
	\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%------------------------------------------------------------------------------



% Pandoc CSL macros
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\begin{document}

% Print the header with above personal informations
% Give optional argument to change alignment(C: center, L: left, R: right)
\makecvheader

% Print the footer with 3 arguments(<left>, <center>, <right>)
% Leave any of these blank if they are not needed
% 2019-02-14 Chris Umphlett - add flexibility to the document name in footer, rather than have it be static Curriculum Vitae
\makecvfooter
  {April 2021}
    {Sarah Gaichas~~~·~~~Curriculum Vitae}
  {\thepage}


%-------------------------------------------------------------------------------
%	CV/RESUME CONTENT
%	Each section is imported separately, open each file in turn to modify content
%------------------------------------------------------------------------------



\hypertarget{education}{%
\section{Education}\label{education}}

\begin{cventries}
    \cventry{PhD}{University of Washington}{Seattle}{2000 - 2006}{}\vspace{-4.0mm}
    \cventry{M.S.}{College of William and Mary, School of Marine Science}{Gloucester Point}{1994 - 1997}{}\vspace{-4.0mm}
    \cventry{B.A.}{Swarthmore College}{Swarthmore}{1986 - 1991}{}\vspace{-4.0mm}
\end{cventries}

\hypertarget{graduate-supervision-and-examination}{%
\section{Graduate Supervision and
Examination}\label{graduate-supervision-and-examination}}

\textbf{Adjunct professor} \emph{University of Massachusetts, Dartmouth}

One current UMassD PhD student committee (April 2021 Exam)\\
One current College of William and Mary (Virginia Institute of Marine
Science) PhD student committee\\
One current UNH PhD student committee

2020 External examiner for a PhD student at University of Iceland\\
2019 (Exam) PhD student committee at the University of Massachusetts,
Dartmouth\\
2019 External examiner for a PhD student at Dalhousie University\\
2016 External examiner for a PhD student at the University of Cape
Town\\
2016 (Exam) PhD student committee at the University of Washington\\
2015 (Exam) PhD student committee at Oregon State University

\hypertarget{experience}{%
\section{Experience}\label{experience}}

\textbf{Research Fishery Biologist} \emph{NOAA/NMFS Northeast Fisheries
Sci. Center, Woods Hole, MA}

\begin{verbatim}
Ecosystem Assessment, MSE       September 2011 - present
\end{verbatim}

Contribute to integrated ecosystem assessments and management strategy
evaluations (MSEs) for the Northeast U.S. continental shelf by analyzing
research survey, fishery, oceanographic, and economic data, and by
developing and testing multispecies assessment models and ecosystem
indicators. Conduct research on ecosystem based management reference
points balancing multiple objectives. Chair National and Northeast MSE
working groups to build capacity and coordinate work plans. Lead editor,
State of the Ecosystem reports. Provide results in peer-reviewed
literature and to the New England and Mid-Atlantic Fishery Management
Councils.

\textbf{Research Fishery Biologist} \emph{NOAA/NMFS Alaska Fisheries
Science Center, Seattle, WA}

\begin{verbatim}
Ecosystem Modeling          June 2006 – September 2011
\end{verbatim}

Analyzed food web, fishery, and research survey data and developed
ecosystem models to evaluate the cumulative effects of fishing, climate
change, and other factors on the marine ecosystem.

\begin{verbatim}
Stock Assessment            November 1998 – June 2006
\end{verbatim}

Assessed the condition of commercially exploited fish stocks through
analysis of fishery and research survey data combined with population
dynamics modeling.

\begin{verbatim}
Observer Program            March 1997 – October 1999
\end{verbatim}

Conducted statistical analyses of fishery observer data, evaluated
present and proposed data collection methods through directed research,
and examined uses of observer data in multiple applications.

\hypertarget{research-science-and-management-organizations}{%
\section{Research, Science, and Management
Organizations}\label{research-science-and-management-organizations}}

\begin{itemize}
\tightlist
\item
  Mid-Atlantic FMC Scientific and Statistical Committee, (2014-present);
  determining Allowable Biological Catch (ABC) advice for fishery
  management and other scientific support.
\item
  Member and past co-chair of the ICES Working Group on North Atlantic
  Regional Seas; developing ecosystem indicators and integrated
  ecosystem assessments for Canadian and U.S. waters.
\item
  Co-Chair of the ICES Working Group on Multispecies Assessment Methods:
  reviewing and contributing to multispecies and ecosystem modeling
  progress across ICES regions.
\item
  Advisory partner in the European Commission's PANDORA project
  (2018-2022), which is designing toolsets for new dynamic ocean
  resource assessments and exploitation.
\end{itemize}

\hypertarget{publications}{%
\section{Publications}\label{publications}}

\protect\hyperlink{cite-bastille_improving_2021}{Bastille, K., S.
Hardison, L. deWitt, J. Brown, J. Samhouri, S. Gaichas, S. Lucey, K.
Kearney, B. Best, S. Cross, S. Large, and E. Spooner} (2021).
``Improving the IEA Approach Using Principles of Open Data Science.''
In: \emph{Coastal Management} 49.1, pp.~72-89. DOI:
\href{https://doi.org/10.1080\%2F08920753.2021.1846155}{10.1080/08920753.2021.1846155}.

\protect\hyperlink{cite-depiper_learning_2021}{DePiper, G., S. Gaichas,
B. Muffley, G. Ardini, J. Brust, J. Coakley, K. Dancy, G. W. Elliott, D.
C. Leaning, D. Lipton, J. McNamee, C. Perretti, K. Rootes-Murdy, and M.
J. Wilberg} (2021). ``Learning by doing: collaborative conceptual
modelling as a path forward in ecosystem-based management.'' In:
\emph{ICES Journal of Marine Science}. DOI:
\href{https://doi.org/10.1093\%2Ficesjms\%2Ffsab054}{10.1093/icesjms/fsab054}.

\protect\hyperlink{cite-lucey_evaluating_2021}{Lucey, S. M., K. Y.
Aydin, S. K. Gaichas, S. X. Cadrin, G. Fay, M. J. Fogarty, and A. Punt}
(2021). ``Evaluating fishery management strategies using an ecosystem
model as an operating model.'' In: \emph{Fisheries Research} 234,
p.~105780. DOI:
\href{https://doi.org/10.1016\%2Fj.fishres.2020.105780}{10.1016/j.fishres.2020.105780}.

\protect\hyperlink{cite-muffley_there_2021}{Muffley, B., S. Gaichas, G.
DePiper, R. Seagraves, and S. Lucey} (2021). ``There Is no I in EAFM
Adapting Integrated Ecosystem Assessment for Mid-Atlantic Fisheries
Management.'' In: \emph{Coastal Management} 49.1, pp. 90-106. DOI:
\href{https://doi.org/10.1080\%2F08920753.2021.1846156}{10.1080/08920753.2021.1846156}.

\protect\hyperlink{cite-reum_its_2021}{Reum, J. C. P., H. Townsend, S.
Gaichas, S. Sagarese, I. C. Kaplan, and A. Grüss} (2021). ``It's Not the
Destination, It's the Journey: Multispecies Model Ensembles for
Ecosystem Approaches to Fisheries Management.'' In: \emph{Frontiers in
Marine Science} 8. DOI:
\href{https://doi.org/10.3389\%2Ffmars.2021.631839}{10.3389/fmars.2021.631839}.

\protect\hyperlink{cite-link_changing_2020}{Link, J. S., G. Huse, S.
Gaichas, and A. R. Marshak} (2020). ``Changing how we approach
fisheries: A first attempt at an operational framework for ecosystem
approaches to fisheries management.'' In: \emph{Fish and Fisheries}
21.2, pp.~393-434. DOI:
\href{https://doi.org/10.1111\%2Ffaf.12438}{10.1111/faf.12438}.

\protect\hyperlink{cite-lucey_conducting_2020}{Lucey, S. M., S. K.
Gaichas, and K. Y. Aydin} (2020). ``Conducting reproducible ecosystem
modeling using the open source mass balance model Rpath.'' In:
\emph{Ecological Modelling} 427, p.~109057. DOI:
\href{https://doi.org/10.1016\%2Fj.ecolmodel.2020.109057}{10.1016/j.ecolmodel.2020.109057}.

\protect\hyperlink{cite-weiskopf_climate_2020}{Weiskopf, S. R., M. A.
Rubenstein, L. G. Crozier, S. Gaichas, R. Griffis, J. E. Halofsky, K. J.
W. Hyde, T. L. Morelli, J. T. Morisette, R. C. Muñoz, A. J. Pershing, D.
L. Peterson, R. Poudel, M. D. Staudinger, A. E. Sutton-Grier, L.
Thompson, J. Vose, J. F. Weltzin, and K. P. Whyte} (2020). ``Climate
change effects on biodiversity, ecosystems, ecosystem services, and
natural resource management in the United States.'' In: \emph{Science of
The Total Environment}, p.~137782. DOI:
\href{https://doi.org/10.1016\%2Fj.scitotenv.2020.137782}{10.1016/j.scitotenv.2020.137782}.

\protect\hyperlink{cite-goethel_recent_2019}{Goethel, D. R., S. M.
Lucey, A. M. Berger, S. K. Gaichas, M. A. Karp, P. D. Lynch, and J. F.
Walter} (2019). ``Recent advances in management strategy evaluation:
introduction to the special issue ``Under pressure: addressing fisheries
challenges with Management Strategy Evaluation''". In: \emph{Canadian
Journal of Fisheries and Aquatic Sciences} 76.10, pp.~1689-1696. DOI:
\href{https://doi.org/10.1139\%2Fcjfas-2019-0084}{10.1139/cjfas-2019-0084}.

\protect\hyperlink{cite-karp_accounting_2019}{Karp, M. A., J. O.
Peterson, P. D. Lynch, R. B. Griffis, C. F. Adams, W. S. Arnold, L. A.
K. Barnett, Y. deReynier, J. DiCosimo, K. H. Fenske, S. K. Gaichas, A.
Hollowed, K. Holsman, M. Karnauskas, D. Kobayashi, A. Leising, J. P.
Manderson, M. McClure, W. E. Morrison, E. Schnettler, A. Thompson, J. T.
Thorson, J. F. Walter, A. J. Yau, R. D. Methot, and J. S. Link} (2019).
``Accounting for shifting distributions and changing productivity in the
development of scientific advice for fishery management.'' In:
\emph{ICES Journal of Marine Science} 76.5, pp.~1305-1315. DOI:
\href{https://doi.org/10.1093\%2Ficesjms\%2Ffsz048}{10.1093/icesjms/fsz048}.

\protect\hyperlink{cite-townsend_progress_2019}{Townsend, H., C. J.
Harvey, Y. deReynier, D. Davis, S. G. Zador, S. Gaichas, M. Weijerman,
E. L. Hazen, and I. C. Kaplan} (2019). ``Progress on Implementing
Ecosystem-Based Fisheries Management in the United States Through the
Use of Ecosystem Models and Analysis.'' In: \emph{Frontiers in Marine
Science} 6. DOI:
\href{https://doi.org/10.3389\%2Ffmars.2019.00641}{10.3389/fmars.2019.00641}.

\protect\hyperlink{cite-deroba_dream_2018}{Deroba, J. J., S. K. Gaichas,
M. Lee, R. G. Feeney, D. V. Boelke, and B. J. Irwin} (2018). ``The dream
and the reality: meeting decision-making time frames while incorporating
ecosystem and economic models into management strategy evaluation.'' In:
\emph{Canadian Journal of Fisheries and Aquatic Sciences}. DOI:
\href{https://doi.org/10.1139\%2Fcjfas-2018-0128}{10.1139/cjfas-2018-0128}.

\protect\hyperlink{cite-feeney_integrating_2018}{Feeney, R. G., D. V.
Boelke, J. J. Deroba, S. K. Gaichas, B. J. Irwin, and M. Lee} (2018).
``Integrating Management Strategy Evaluation into fisheries management:
advancing best practices for stakeholder inclusion based on an MSE for
Northeast U.S. Atlantic herring.'' In: \emph{Canadian Journal of
Fisheries and Aquatic Sciences}. DOI:
\href{https://doi.org/10.1139\%2Fcjfas-2018-0125}{10.1139/cjfas-2018-0125}.

\protect\hyperlink{cite-gaichas_implementing_2018}{Gaichas, S. K., G. S.
DePiper, R. J. Seagraves, B. W. Muffley, M. Sabo, L. L. Colburn, and A.
L. Loftus} (2018). ``Implementing Ecosystem Approaches to Fishery
Management: Risk Assessment in the US Mid-Atlantic.'' In:
\emph{Frontiers in Marine Science} 5. DOI:
\href{https://doi.org/10.3389\%2Ffmars.2018.00442}{10.3389/fmars.2018.00442}.

\protect\hyperlink{cite-goethel_closing_2018}{Goethel, D. R., S. M.
Lucey, A. M. Berger, S. K. Gaichas, M. A. Karp, P. D. Lynch, J. F.
Walter III, J. J. Deroba, S. K. Miller, and M. J. Wilberg} (2018).
``Closing the Feedback Loop: On Stakeholder Participation in Management
Strategy Evaluation.'' In: \emph{Canadian Journal of Fisheries and
Aquatic Sciences}. DOI:
\href{https://doi.org/10.1139\%2Fcjfas-2018-0162}{10.1139/cjfas-2018-0162}.

\protect\hyperlink{cite-lipton_ecosystems_2018}{Lipton, D., M. A.
Rubenstein, S. R. Weiskopf, S. L. Carter, J. Peterson, L. Crozier, M.
Fogarty, S. Gaichas, K. J. W. Hyde, T. L. Morelli, J. Morisette, H.
Moustahfid, R. Munoz, R. Poudel, M. D. Staudinger, C. Stock, L.
Thompson, R. S. Waples, and J. F. Weltzin} (2018). \emph{Ecosystems,
Ecosystem Services, and Biodiversity}. Other Government Series. U.S.
Global Change Research Program, p.~268321. DOI:
\href{https://doi.org/10.7930\%2FNCA4.2018.CH7}{10.7930/NCA4.2018.CH7}.

\protect\hyperlink{cite-olsen_ocean_2018}{Olsen, E., I. C. Kaplan, C.
Ainsworth, G. Fay, S. Gaichas, R. Gamble, R. Girardin, C. H. Eide, T. F.
Ihde, H. N. Morzaria-Luna, K. F. Johnson, M. Savina-Rolland, H.
Townsend, M. Weijerman, E. A. Fulton, and J. S. Link} (2018). ``Ocean
Futures Under Ocean Acidification, Marine Protection, and Changing
Fishing Pressures Explored Using a Worldwide Suite of Ecosystem
Models.'' In: \emph{Frontiers in Marine Science} 5. DOI:
\href{https://doi.org/10.3389\%2Ffmars.2018.00064}{10.3389/fmars.2018.00064}.

\protect\hyperlink{cite-depiper_operationalizing_2017}{DePiper, G. S.,
S. K. Gaichas, S. M. Lucey, P. Pinto da Silva, M. R. Anderson, H.
Breeze, A. Bundy, P. M. Clay, G. Fay, R. J. Gamble, R. S. Gregory, P. S.
Fratantoni, C. L. Johnson, M. Koen-Alonso, K. M. Kleisner, J. Olson, C.
T. Perretti, P. Pepin, F. Phelan, V. S. Saba, L. A. Smith, J. C. Tam, N.
D. Templeman, and R. P. Wildermuth} (2017). ``Operationalizing
integrated ecosystem assessments within a multidisciplinary team:
lessons learned from a worked example.'' In: \emph{ICES Journal of
Marine Science} 74.8, pp.~2076-2086. DOI:
\href{https://doi.org/10.1093\%2Ficesjms\%2Ffsx038}{10.1093/icesjms/fsx038}.

\protect\hyperlink{cite-gaichas_combining_2017}{Gaichas, S. K., M.
Fogarty, G. Fay, R. Gamble, S. Lucey, and L. Smith} (2017). ``Combining
stock, multispecies, and ecosystem level fishery objectives within an
operational management procedure: simulations to start the
conversation.'' In: \emph{ICES Journal of Marine Science} 74.2, pp.
552-565. DOI:
\href{https://doi.org/10.1093\%2Ficesjms\%2Ffsw119}{10.1093/icesjms/fsw119}.

\protect\hyperlink{cite-holsman_ecosystem-based_2017}{Holsman, K., J.
Samhouri, G. Cook, E. Hazen, E. Olsen, M. Dillard, S. Kasperski, S.
Gaichas, C. R. Kelble, M. Fogarty, and K. Andrews} (2017). ``An
ecosystem-based approach to marine risk assessment.'' In:
\emph{Ecosystem Health and Sustainability} 3.1, p.~e01256. DOI:
\href{https://doi.org/10.1002\%2Fehs2.1256}{10.1002/ehs2.1256}.

\protect\hyperlink{cite-tommasi_managing_2017}{Tommasi, D., C. A. Stock,
A. J. Hobday, R. Methot, I. C. Kaplan, J. P. Eveson, K. Holsman, T. J.
Miller, S. Gaichas, M. Gehlen, A. Pershing, G. A. Vecchi, R. Msadek, T.
Delworth, C. M. Eakin, M. A. Haltuch, R. Séférian, C. M. Spillman, J. R.
Hartog, S. Siedlecki, J. F. Samhouri, B. Muhling, R. G. Asch, M. L.
Pinsky, V. S. Saba, S. B. Kapnick, C. F. Gaitan, R. R. Rykaczewski, M.
A. Alexander, Y. Xue, K. V. Pegion, P. Lynch, M. R. Payne, T.
Kristiansen, P. Lehodey, and F. E. Werner} (2017). ``Managing living
marine resources in a dynamic environment: The role of seasonal to
decadal climate forecasts.'' In: \emph{Progress in Oceanography} 152,
pp.~15-49. DOI:
\href{https://doi.org/10.1016\%2Fj.pocean.2016.12.011}{10.1016/j.pocean.2016.12.011}.

\protect\hyperlink{cite-wildermuth_structural_2017}{Wildermuth, R. P.,
G. Fay, and S. Gaichas} (2017). ``Structural uncertainty in qualitative
models for ecosystem-based management of Georges Bank.'' In:
\emph{Canadian Journal of Fisheries and Aquatic Sciences} 75.10,
pp.~1635-1643. DOI:
\href{https://doi.org/10.1139\%2Fcjfas-2017-0149}{10.1139/cjfas-2017-0149}.

\protect\hyperlink{cite-zador_linking_2017}{Zador, S. G., S. K. Gaichas,
S. Kasperski, C. L. Ward, R. E. Blake, N. C. Ban, A. Himes-Cornell, and
J. Z. Koehn} (2017). ``Linking ecosystem processes to communities of
practice through commercially fished species in the Gulf of Alaska.''
In: \emph{ICES Journal of Marine Science} 74.7, pp.~2024-2033. DOI:
\href{https://doi.org/10.1093\%2Ficesjms\%2Ffsx054}{10.1093/icesjms/fsx054}.

\protect\hyperlink{cite-gaichas_framework_2016}{Gaichas, S. K., R. J.
Seagraves, J. M. Coakley, G. S. DePiper, V. G. Guida, J. A. Hare, P. J.
Rago, and M. J. Wilberg} (2016). ``A Framework for Incorporating
Species, Fleet, Habitat, and Climate Interactions into Fishery
Management.'' In: \emph{Frontiers in Marine Science} 3. DOI:
\href{https://doi.org/10.3389\%2Ffmars.2016.00105}{10.3389/fmars.2016.00105}.

\protect\hyperlink{cite-olsen_ecosystem_2016}{Olsen, E., G. Fay, S.
Gaichas, R. Gamble, S. Lucey, and J. S. Link} (2016). ``Ecosystem Model
Skill Assessment. Yes We Can!'' In: \emph{PLOS ONE} 11.1. Ed. by C. N.
Bianchi, p.~e0146467. DOI:
\href{https://doi.org/10.1371\%2Fjournal.pone.0146467}{10.1371/journal.pone.0146467}.

\protect\hyperlink{cite-weijerman_atlantis_2016}{Weijerman, M., J. Link,
E. Fulton, E. Olsen, H. Townsend, S. Gaichas, C. Hansen, M.
Skern-Mauritzen, I. Kaplan, R. Gamble, G. Fay, M. Savina, C. Ainsworth,
I. van Putten, R. Gorton, R. Brainard, K. Larsen, and T. Hutton} (2016).
``Atlantis Ecosystem Model Summit: Report from a workshop.'' In:
\emph{Ecological Modelling} 335, pp.~35-38. DOI:
\href{https://doi.org/10.1016\%2Fj.ecolmodel.2016.05.007}{10.1016/j.ecolmodel.2016.05.007}.

\protect\hyperlink{cite-zador_ecosystem_2016}{Zador, S. G., K. K.
Holsman, K. Y. Aydin, and S. K. Gaichas} (2016). ``Ecosystem
considerations in Alaska: the value of qualitative assessments.'' In:
\emph{ICES Journal of Marine Science: Journal du Conseil}, p.~fsw144.
DOI:
\href{https://doi.org/10.1093\%2Ficesjms\%2Ffsw144}{10.1093/icesjms/fsw144}.

\protect\hyperlink{cite-gaichas_wasp_2015}{Gaichas, S., K. Aydin, and R.
C. Francis} (2015). ``Wasp waist or beer belly? Modeling food web
structure and energetic control in Alaskan marine ecosystems, with
implications for fishing and environmental forcing.'' In: \emph{Progress
in Oceanography} 138, Part A, pp.~1-17. DOI:
\href{https://doi.org/10.1016\%2Fj.pocean.2015.09.010}{10.1016/j.pocean.2015.09.010}.

\protect\hyperlink{cite-smith_simulations_2015}{Smith, L., R. Gamble, S.
Gaichas, and J. Link} (2015). ``Simulations to evaluate management
trade-offs among marine mammal consumption needs, commercial fishing
fleets and finfish biomass.'' In: \emph{Marine Ecology Progress Series}
523, pp.~215-232. DOI:
\href{https://doi.org/10.3354\%2Fmeps11129}{10.3354/meps11129}.

\protect\hyperlink{cite-gaichas_risk-based_2014}{Gaichas, S. K., J. S.
Link, and J. A. Hare} (2014). ``A risk-based approach to evaluating
northeast US fish community vulnerability to climate change.'' In:
\emph{ICES Journal of Marine Science} 71.8, pp. 2323-2342. DOI:
\href{https://doi.org/10.1093\%2Ficesjms\%2Ffsu048}{10.1093/icesjms/fsu048}.

\protect\hyperlink{cite-ruzicka_dividing_2013}{Ruzicka, J. J., J. H.
Steele, T. Ballerini, S. K. Gaichas, and D. G. Ainley} (2013).
``Dividing up the pie: Whales, fish, and humans as competitors.'' In:
\emph{Progress in Oceanography} 116, pp.~207-219. DOI:
\href{https://doi.org/10.1016\%2Fj.pocean.2013.07.009}{10.1016/j.pocean.2013.07.009}.

\protect\hyperlink{cite-ruzicka_analysis_2013}{Ruzicka, J., J. Steele,
S. Gaichas, T. Ballerini, D. Gifford, R. Brodeur, and E. Hofmann}
(2013). ``Analysis of Energy Flow in US GLOBEC Ecosystems Using
End-to-End Models.'' In: \emph{Oceanography} 26.4, pp.~82-97. DOI:
\href{https://doi.org/10.5670\%2Foceanog.2013.77}{10.5670/oceanog.2013.77}.

\protect\hyperlink{cite-fu_relative_2012}{Fu, C., S. Gaichas, J. Link,
A. Bundy, J. Boldt, A. Cook, R. Gamble, K. Rong Utne, H. Liu, and K.
Friedland} (2012). ``Relative importance of fisheries, trophodynamic and
environmental drivers in a series of marine ecosystems.'' In:
\emph{Marine Ecology Progress Series} 459, pp. 169-184. DOI:
\href{https://doi.org/10.3354\%2Fmeps09805}{10.3354/meps09805}.

\protect\hyperlink{cite-gaichas_beyond_2012}{Gaichas, S. K., G. Odell,
K. Y. Aydin, and R. C. Francis} (2012). ``Beyond the defaults:
functional response parameter space and ecosystem-level fishing
thresholds in dynamic food web model simulations.'' In: \emph{Canadian
Journal of Fisheries and Aquatic Sciences} 69, pp.~2077-2094. DOI:
\href{https://doi.org/10.1139\%2FF2012-099}{10.1139/F2012-099}.

\protect\hyperlink{cite-gaichas_what_2012}{Gaichas, S., A. Bundy, T.
Miller, E. Moksness, and K. Stergiou} (2012). ``What drives marine
fisheries production?'' In: \emph{Marine Ecology Progress Series} 459,
pp.~159-163. DOI:
\href{https://doi.org/10.3354\%2Fmeps09841}{10.3354/meps09841}.

\protect\hyperlink{cite-gaichas_assembly_2012}{Gaichas, S., R. Gamble,
M. J. Fogarty, H. Benoit, T. Essington, C. Fu, M. Koen-Alonso, and J.
Link} (2012). ``Assembly Rules for Aggregate-Species Production Models:
Simulations in Support of Management Strategy Evaluation.'' In:
\emph{Marine Ecology Progress Series} 459, pp.~275-292.

\protect\hyperlink{cite-link_synthesizing_2012}{Link, J., S. Gaichas, T.
Miller, T. Essington, A. Bundy, J. Boldt, K. Drinkwater, and E.
Moksness} (2012). ``Synthesizing Lessons Learned from Comparing
Fisheries Production in 13 Northern Hemisphere Ecosystems: Emergent
Fundamental Features.'' In: \emph{Marine Ecology Progress Series} 459,
pp.~293-302.

\protect\hyperlink{cite-link_dealing_2012}{Link, J., T. Ihde, C. Harvey,
S. Gaichas, J. Field, J. Brodziak, H. Townsend, and R. Peterman} (2012).
``Dealing with uncertainty in ecosystem models: The paradox of use for
living marine resource management.'' In: \emph{Progress in Oceanography}
102, pp.~102-114. DOI:
\href{https://doi.org/10.1016\%2Fj.pocean.2012.03.008}{10.1016/j.pocean.2012.03.008}.

\protect\hyperlink{cite-moksness_bernard_2012}{Moksness, E., J. Link, K.
Drinkwater, and S. Gaichas} (2012). ``Bernard Megrey: pioneer of
Comparative Marine Ecosystem analyses.'' In: \emph{Marine Ecology
Progress Series} 459, pp.~165-167. DOI:
\href{https://doi.org/10.3354\%2Fmeps09582}{10.3354/meps09582}.

\protect\hyperlink{cite-pranovi_trophic-level_2012}{Pranovi, F., J.
Link, C. Fu, A. Cook, H. Liu, S. Gaichas, K. Friedland, K. Rong Utne,
and H. Benoît} (2012). ``Trophic-level determinants of biomass
accumulation in marine ecosystems.'' In: \emph{Marine Ecology Progress
Series} 459, pp.~185-201. DOI:
\href{https://doi.org/10.3354\%2Fmeps09738}{10.3354/meps09738}.

\protect\hyperlink{cite-gaichas_what_2011}{Gaichas, S. K., K. Y. Aydin,
and R. C. Francis} (2011). ``What drives dynamics in the Gulf of Alaska?
Integrating hypotheses of species, fishing, and climate relationships
using ecosystem modeling.'' In: \emph{Canadian Journal of Fisheries and
Aquatic Sciences} 68, pp.~1553-1578.

\protect\hyperlink{cite-hunsicker_functional_2011}{Hunsicker, M. E., L.
Ciannelli, K. M. Bailey, J. A. Buckel, J. W. White, J. S. Link, T. E.
Essington, S. Gaichas, T. W. Anderson, R. D. Brodeur, K. Chan, K. Chen,
G. Englund, K. T. Frank, V. Freitas, M. A. Hixon, T. Hurst, D. W.
Johnson, J. F. Kitchell, D. Reese, G. A. Rose, H. Sjodin, W. J. Sydeman,
H. W. v. d.~Veer, K. Vollset, and S. Zador} (2011). ``Functional
responses and scaling in predator--prey interactions of marine fishes:
contemporary issues and emerging concepts.'' In: \emph{Ecology Letters}
14.12, pp.~1288-1299. DOI:
\href{https://doi.org/10.1111\%2Fj.1461-0248.2011.01696.x\%4010.1111\%2F\%28ISSN\%291461-0248.anthropogenic-change}{10.1111/j.1461-0248.2011.01696.x@10.1111/(ISSN)1461-0248.anthropogenic-change}.

\protect\hyperlink{cite-salomon_bridging_2011}{Salomon, A. K., S. K.
Gaichas, O. P. Jensen, V. N. Agostini, N. Sloan, J. Rice, T. R.
McClanahan, M. H. Ruckelshaus, P. S. Levin, N. K. Dulvy, and E. A.
Babcock} (2011). ``Bridging the Divide Between Fisheries and Marine
Conservation Science.'' In: \emph{Bulletin of Marine Science} 87.2,
pp.~251-274. DOI:
\href{https://doi.org/10.5343\%2Fbms.2010.1089}{10.5343/bms.2010.1089}.

\protect\hyperlink{cite-gaichas_using_2010}{Gaichas, S. K., K. Y. Aydin,
and R. C. Francis} (2010). ``Using food web model results to inform
stock assessment estimates of mortality and production for
ecosystem-based fisheries management.'' In: \emph{Canadian Journal of
Fisheries and Aquatic Sciences} 67, pp.~1493-1506.

\protect\hyperlink{cite-reuter_managing_2010}{Reuter, R. F., M. E.
Conners, J. Dicosimo, S. Gaichas, O. Ormseth, and T. T. Tenbrink}
(2010). ``Managing non-target, data-poor species using catch limits:
lessons from the Alaskan groundfish fishery.'' In: \emph{Fisheries
Management and Ecology} 17.4, pp. 323-335. DOI:
\href{https://doi.org/10.1111\%2Fj.1365-2400.2009.00726.x}{10.1111/j.1365-2400.2009.00726.x}.

\protect\hyperlink{cite-salomon_key_2010}{Salomon, A. K., S. K. Gaichas,
N. T. Shears, J. E. Smith, E. M. P. Madin, and S. D. Gaines} (2010).
``Key Features and Context-Dependence of Fishery-Induced Trophic
Cascades.'' In: \emph{Conservation Biology} 24.2, pp.~382-394. DOI:
\href{https://doi.org/10.1111\%2Fj.1523-1739.2009.01436.x}{10.1111/j.1523-1739.2009.01436.x}.

\protect\hyperlink{cite-gaichas_comparison_2009}{Gaichas, S., G. Skaret,
J. Falk-Petersen, J. S. Link, W. Overholtz, B. A. Megrey, H. Gjøsæter,
W. T. Stockhausen, A. Dommasnes, K. D. Friedland, and K. Aydin} (2009).
``A comparison of community and trophic structure in five marine
ecosystems based on energy budgets and system metrics.'' In:
\emph{Progress in Oceanography} 81.1-4, pp.~47-62. DOI:
\href{https://doi.org/10.1016\%2Fj.pocean.2009.04.005}{10.1016/j.pocean.2009.04.005}.

\protect\hyperlink{cite-link_comparison_2009}{Link, J. S., W. T.
Stockhausen, G. Skaret, W. Overholtz, B. A. Megrey, H. Gjøsæter, S.
Gaichas, A. Dommasnes, J. Falk-Petersen, J. Kane, F. J. Mueter, K. D.
Friedland, and J. A. Hare} (2009). ``A comparison of biological trends
from four marine ecosystems: Synchronies, differences, and
commonalities.'' In: \emph{Progress in Oceanography}. Comparative Marine
Ecosystem Structure and Function: Descriptors and Characteristics 81.1,
pp.~29-46. DOI:
\href{https://doi.org/10.1016\%2Fj.pocean.2009.04.004}{10.1016/j.pocean.2009.04.004}.

\protect\hyperlink{cite-megrey_cross-ecosystem_2009}{Megrey, B. A., J.
A. Hare, W. T. Stockhausen, A. Dommasnes, H. Gjøsæter, W. Overholtz, S.
Gaichas, G. Skaret, J. Falk-Petersen, J. S. Link, and K. D. Friedland}
(2009). ``A cross-ecosystem comparison of spatial and temporal patterns
of covariation in the recruitment of functionally analogous fish
stocks.'' In: \emph{Progress in Oceanography}. Comparative Marine
Ecosystem Structure and Function: Descriptors and Characteristics 81.1,
pp.~63-92. DOI:
\href{https://doi.org/10.1016\%2Fj.pocean.2009.04.006}{10.1016/j.pocean.2009.04.006}.

\protect\hyperlink{cite-gaichas_context_2008}{Gaichas, S.} (2008). ``A
context for ecosystem-based fishery management: Developing concepts of
ecosystems and sustainability.'' In: \emph{Marine Policy} 32,
pp.~393-401.

\protect\hyperlink{cite-gaichas_network_2008}{Gaichas, S. K. and R. C.
Francis} (2008). ``Network models for ecosystem-based fishery analysis:
A review of concepts and application to the Gulf of Alaska marine food
web.'' In: \emph{Canadian Journal of Fisheries and Aquatic Sciences} 65,
pp.~1965-1982.

\protect\hyperlink{cite-aydin_comparison_2007}{Aydin, K. Y., S. Gaichas,
I. Ortiz, D. Kinzey, and N. Friday} (2007). \emph{A comparison of the
Bering Sea, Gulf of Alaska, and Aleutian Islands large marine ecosystems
through food web modeling}. U.S. Department of Commerce, NOAA Tech.
Memo. NMFS-AFSC-178, p.~298.

\protect\hyperlink{cite-gburski_age_2007}{Gburski, C. M., S. K. Gaichas,
and D. K. Kimura} (2007). ``Age and growth of big skate (Raja
binoculata) and longnose skate (R. rhina) in the Gulf of Alaska.'' In:
\emph{Environmental Biology of Fishes} 80.2-3, pp.~337-349. DOI:
\href{https://doi.org/10.1007\%2Fs10641-007-9231-8}{10.1007/s10641-007-9231-8}.

\protect\hyperlink{cite-gaichas_development_2006}{Gaichas, S.} (2006).
``Development and application of ecosystem models to support fishery
sustainability: A case study for the Gulf of Alaska - ProQuest.''
Seattle, WA.

\protect\hyperlink{cite-spies_dna-based_2006}{Spies, I. B., S. Gaichas,
D. E. Stevenson, J. W. Orr, and M. F. Canino} (2006). ``DNA-based
identification of Alaska skates (Amblyraja, Bathyraja and Raja: Rajidae)
using cytochrome c oxidase subunit I (coI) variation.'' In:
\emph{Journal of Fish Biology} 69.sb, pp.~283-292. DOI:
\href{https://doi.org/10.1111\%2Fj.1095-8649.2006.01286.x}{10.1111/j.1095-8649.2006.01286.x}.

\protect\hyperlink{cite-barbeaux_evaluation_2005}{Barbeaux, S. J., S.
Gaichas, J. N. Ianelli, and M. W. Dorn} (2005). ``Evaluation of
Biological Sampling Protocols for At-Sea Groundfish Observers in
Alaska.'' In: \emph{Alaska Fishery Research Bulletin} 11.2, p. 25.

\protect\hyperlink{cite-karp_government-industry_2001}{Karp, W. A., C.
S. Rose, J. R. Gauvin, S. K. Gaichas, M. W. Dorn, and G. D. Stauffer}
(2001). ``Government-Industry Cooperative Fisheries Research in the
North Pacific under the MSFCMA.'' In: \emph{Marine Fisheries Review}
63.1, pp. 40-46.

\protect\hyperlink{cite-dorn_measuring_1999}{Dorn, M. W., S. K. Gaichas,
S. M. Fitzgerald, and S. A. Bibb} (1999). ``Measuring Total Catch at
Sea: Use of a Motion-Compensated Flow Scale to Evaluate Observer
Volumetric Methods.'' In: \emph{North American Journal of Fisheries
Management} 19.4, pp.~999-1016. DOI:
\href{https://doi.org/10.1577\%2F1548-8675\%281999\%29019\%3C0999\%3AMTCASU\%3E2.0.CO\%3B2}{10.1577/1548-8675(1999)019\textless0999:MTCASU\textgreater2.0.CO;2}.

\protect\hyperlink{cite-gaichas_age_1997}{Gaichas, S.} (1997). ``Age And
Growth Of Spanish Mackerel, Scomberomorus Maculatus, In The Chesapeake
Bay Region.'' In: \emph{Master's Thesis}. DOI:
\href{https://doi.org/10.25773\%2FV5-TWBW-6T04}{10.25773/V5-TWBW-6T04}.

\hypertarget{scientific-presentations}{%
\section{Scientific Presentations}\label{scientific-presentations}}

Speaker, moderator, and symposium organizer at regional to international
professional meetings (AFS, ASLO, Conservation Biology, ESA, GLOBEC,
IMCC, World Fisheries Congress) 1995-present. Invited speaker for public
lecture series, teacher workshops, and university courses and
departmental seminars.

\end{document}
