---
address: |
  NOAA NMFS Northeast Fisheries Science Center
  166 Water Street, Woods Hole, MA 02543
date: "`r format(Sys.time(), '%B %Y')`"
email: Sarah.Gaichas@noaa.gov
github: sgaichas
name: Sarah
output:
  vitae::awesomecv:
    keep_tex: yes
  vitae::moderncv: default
phone: +1 508 495 2016
position: Research Fishery Biologist ZP-IV
surname: Gaichas
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
```

# Educational History

```{r}
library(tibble)

#rorcid::orcid_auth()

orcid_data <- do.call("rbind",
    rorcid::orcid_educations("0000-0002-5788-3073")$`0000-0002-5788-3073`$`affiliation-group`$summaries
)

orcid_data %>%
  detailed_entries(
    what = `education-summary.role-title`,
    when = glue::glue("{`education-summary.start-date.year.value`} - {`education-summary.end-date.year.value`}"),
    with = `education-summary.organization.name`,
    where = glue::glue("{`education-summary.organization.address.city`},{`education-summary.organization.address.region`}"),
  )

# tribble(
#   ~ Degree, ~ Year, ~ Institution, ~ Where,
#   "Informal studies", "1889-91", "Flying University", "Warsaw, Poland",
#   "Master of Physics", "1893", "Sorbonne Université", "Paris, France",
#   "Master of Mathematics", "1894", "Sorbonne Université", "Paris, France"
# ) %>% 
#   detailed_entries(Degree, Year, Institution, Where)

```


# Professional History

**Research Fishery Biologist**                   		*NOAA/NMFS Northeast Fisheries Sci. Center, Woods Hole, MA*

	Ecosystem Assessment, MSE		September 2011 - present
Contribute to integrated ecosystem assessments and management strategy evaluations (MSEs) for the Northeast U.S. continental shelf by analyzing research survey, fishery, oceanographic, and economic data, and by developing and testing multispecies assessment models and ecosystem indicators. Conduct research on ecosystem based management reference points balancing multiple objectives. Chair National and Northeast MSE working groups to build capacity and coordinate work plans. Lead editor, State of the Ecosystem reports. Provide results in peer-reviewed literature and to the New England and Mid-Atlantic Fishery Management Councils. 

**Research Fishery Biologist**                   		*NOAA/NMFS Alaska Fisheries Science Center, Seattle, WA*

	Ecosystem Modeling			June 2006 – September 2011
Established an ecosystem context for fishery management through analysis of food web, fishery, and research survey data combined with ecosystem modeling.  Evaluated the cumulative effects of fishing, climate change, and other factors on the marine ecosystem and species within it. Conducted independent research on ecosystem level indicators and thresholds and provided results in peer-reviewed literature and to the North Pacific Fishery Management Council. Participated in research surveys at sea. 

	Stock Assessment			November 1998 – June 2006
Assessed the condition of commercially exploited fish stocks through analysis of fishery and research survey data combined with population dynamics modeling. Evaluated the effects of fishing on non-target species and the marine ecosystem.  Conducted Ph.D. research. Attended 2001 Complex Systems Summer School, Santa Fe Institute, Santa Fe NM. Participated in research surveys at sea.

Stock assessments:  1999 Gulf of Alaska (GOA) and 2000 – 2004 Bering Sea Aleutian Islands (BSAI) “Other species” (all species of skates, sharks, sculpins, octopi, and squids), 2000-2005 BSAI squids, 2005 BSAI skates, 2003-2005 GOA skates, 2000-2005 GOA Thornyheads, and 2006 GOA squids.  


	Observer Program			March 1997 – October 1999
Conducted statistical analyses of fishery observer data, evaluated present and proposed data collection methods through directed research, and examined uses of observer data in multiple applications. 

# Awards

* End-year performance bonuses each year, 2006-2020.
* Individual cash award for State of the Ecosystem reporting, August 2020.
* Individual cash award for Acting Branch Chief performance, July 2017.
* ICES/MYFISH symposium Targets and Limits for Long Term Fisheries Management Most Inspiring Paper, October 2015. 
* Department of Commerce “Cash-In-A-Flash” award, December 2011.
* Department of Commerce Special Act Award: Alaska Fisheries Science Center Employee of the Year, 2010.
* Department of Commerce Special Act Award for service on North Pacific Council Gulf of Alaska Plan Team and for ecosystem research contributions, July 2005.
* Department of Commerce Special Act Award for innovations in nontarget species management and ecosystem modeling, 2004.
* National Association of Environmental Professionals Award for Environmental Excellence in NEPA for assistance with the Steller Seal Lion Protection Measures SEIS, June 2003.
* Mote International Symposium on Confronting Trade-offs in the Ecosystem Approach to Fisheries Management Young Investigator Award, November 2002.
* Department of Commerce Special Act Award for service as the Nontarget and Prohibited Species Team Leader in the development of the Draft Programmatic SEIS for groundfish fisheries in Alaska, September 2000. 
* Department of Commerce “Cash-In-A-Flash” awards, March and June 1998. 
* College of William and Mary William J. Hargis Jr. Fellowship Award and Dean's Fellowship. April, 1995.

# Publications

## Papers published in peer-reviewed journals

```{r bib, warning=FALSE, results='asis'}

if(!file.exists("gaichas.bib")){
scholar::get_publications("C_gOAZ4AAAAJ&hl") %>% 
  #filter(year > 2014) %>% 
  arrange(desc(year)) %>%
  distinct(title, .keep_all = TRUE) %>%
  transmute(bibtype = "Article", author = as.character(author),
            title = as.character(title),
            journaltitle = as.character(journal), year, key = row_number()) %>% 
  RefManageR::as.BibEntry() %>% 
  RefManageR::WriteBib("gaichas.bib")
}

# this bib from zotero; google scholar brought in conference papers, etc
bib <- RefManageR::ReadBib("gaichas.bib")
#sort(bib, sorting = "ydnt")
RefManageR::BibOptions(sorting = "ydnt", 
                       bib.style = "authoryear", 
                       max.names = 50,
                       style = "markdown",
                       no.print.fields = c("abstract", "copyright", "issn", "url", "language","urldate","month","file", "shorttitle", "note", "keywords"))
RefManageR::NoCite(bib)

newline <- list(rep("  \n", length(bib)))
RefManageR::BibEntry()
#bib <- mapply(append, RefManageR::BibEntry(bibtype = "doi"), newline, SIMPLIFY = FALSE)
RefManageR::PrintBibliography(bib)

#bibliography_entries("gaichas.bib")%>%
#  select(author, year, title, journal, pages, url) %>%
#  arrange(desc(year, author$family))
```

## Papers submitted to peer-reviewed journals

Staudinger, M., Lynch, A.,  Gaichas, S., Fox, M., Gibson-Reinemer, D., Langan, J., Teffer, A., Thackeray, S., and Winfield, I.  Accepted April 2021. How does climate change affect emergent properties of aquatic ecosystems? Fisheries.

Kaplan, I.C., S.K. Gaichas, C.C. Stawitz, P.D. Lynch, K.N. Marshall, J.J. Deroba, M. Masi, J.K.T. Brodziak, K.Y. Aydin, K. Holsman, H. Townsend, D. Tommasi, J.A. Smith, S. Koenigstein, M. Weijerman and J. Link. Accepted May 2021. Management Strategy Evaluation: Allowing the Light on the Hill to Illuminate More than One Species. Original Research, Front. Mar. Sci. - Marine Fisheries, Aquaculture and Living Resources. 

## Other papers (book chapters)

Gaichas, S.K., C. Reiss, and M. Koen-Alonso. 2014. Ecosystem Based Management in high latitude ecosystems. Chapter 10, p. 277-324 in The Sea, Volume 16: Marine Ecosystem Based Management. Harvard U. Press.

## Reports

*Northeast Fisheries Science Center State of the Ecosystem Reports; lead editor for Mid-Atlantic 2017-2021, lead editor for New England 2017-2020, supporting editor for New England 2021* 

2021 State of the Ecosystem: 

*  Mid-Atlantic https://apps-nefsc.fisheries.noaa.gov/rcb/publications/SOE-MAFMC-2021-508-Final.pdf
*  New England https://apps-nefsc.fisheries.noaa.gov/rcb/publications/SOE-MAFMC-2021-508-Final.pdf  

2020 State of the Ecosystem: 

*  Mid-Atlantic https://doi.org/10.25923/1f8j-d564
*  New England https://doi.org/10.25923/4tdk-eg57  

2019 State of the Ecosystem:  

*  Mid-Atlantic https://www.mafmc.org/s/Tab09_State-of-the-Ecosystem-Report_2019-04.pdf
*  New England https://s3.amazonaws.com/nefmc.org/11_SOE-NEFMC-2019.pdf  

2018 State of the Ecosystem:  

*  Mid-Atlantic https://www.mafmc.org/s/Tab04_State-of-the-Ecosystem-Apr2018.pdf
*  New England https://s3.amazonaws.com/nefmc.org/2_Ecosystem-Status-Report.pdf  

2017 State of the Ecosystem:  

*  Mid-Atlantic https://www.mafmc.org/s/Tab02_2017-04_State-of-the-Ecosystem-and-EAFM.pdf
*  New England: https://s3.amazonaws.com/nefmc.org/2_2016-State-of-the-Ecosystem-Report.pdf  

*Recent International Council for the Exploration of the Seas (ICES) Expert Group reports*

ICES. 2021. Working Group on Multispecies Assessment Methods (WGSAM; outputs from 2020 meeting). ICES Scientific Reports. 3:10. 231 pp. https://doi.org/10.17895/ices.pub.7695

ICES. 2021. Workshop of Fisheries Management Reference Points in a Changing Environment (WKRPChange, outputs from 2020 meeting). ICES Scientific Reports. 3:6. 39 pp.https://doi.org/10.17895/ices.pub.7660

ICES. 2019. Working Group on Multispecies Assessment Methods (WGSAM). ICES Scientific Reports.1:91. 320 pp. http://doi.org/10.17895/ices.pub.5758

ICES. 2018. Report of the Working Group on Multispecies Assessment Methods (WGSAM), 15–19 October 2018, Paris, France. ICES CM 2018/HAPISG:20. 89 pp.

*Selectied NOAA Technical Memoranda*

Townsend, H., K. Aydin, K. Holsman, C. Harvey, I. Kaplan, E. Hazen, P. Woodworth-Jefcoats, M. Weijerman, T. Kellison, S. Gaichas, K. Osgood, J. Link (editors). 2017. Report of the 4th National Ecosystem Modeling Workshop (NEMoW 4): Using Ecosystem Models to Evaluate Inevitable Trade-offs. U.S. Dept. of Commerce, NOAA Tech. Memo. NMFS-F/SPO-173, 77 p.

Link, J.S., D. Mason, T. Lederhouse, S. Gaichas, T. Hartley, J. Ianelli, R. Methot, C. Stock, C. Stow, and H. Townsend. 2015. Report from the Joint OAR-NMFS Modeling Uncertainty Workshop. U.S. Dept. of Commer., NOAA. NOAA Technical Memorandum NMFS-F/SPO-153, 31 p.

Link, J. S., T.F. Ihde, H. Townsend, K. Osgood, M. Schirripa, D. Kobayashi, S. Gaichas, J. Field, P. Levin, K. Aydin, G. Watters, and C. Harvey (editors). 2010. Report of the 2nd National Ecosystem Modeling Workshop (NEMoW II):  Bridging The Credibility Gap – Dealing With Uncertainty In Ecosystem Models. U.S. Dep. Commerce, NOAA Tech. Memo. NMFS-F/SPO-102, 72 p.

Conners, M.E., J. Cahalan, S. Gaichas, W.A. Karp, T. Loomis, and J. Watson. 2009. Sampling for estimation of catch composition in Bering Sea trawl fisheries. NOAA Tech. Memo. NMFS-AFSC-199, 42 p. 

Aydin, K., S. Gaichas, I. Ortiz, D. Kinzey, and N. Friday. 2007. A comparison of the Bering Sea, Gulf of Alaska, and Aleutian Islands large marine ecosystems through food web modeling. U.S. Dep. Commer., NOAA Tech. Memo. NMFS-AFSC-178, 298 p.

Gaichas, S.K., and J.C. Field, 2003. Comparative modeling approaches for exploited marine ecosystems in the North Pacific: what are these models good for? In Mace, P.M. (ed.), Proceedings of the seventh NMFS National Stock Assessment Workshop, (Re)building Sustainable Fisheries and Marine Ecosystems.  NOAA Tech. Memo NMFS-F/SPO-62.


## Others (Fishery Management Council Documents)

April 2020-2021: Lead author, State of the Ecosystem Request tracking memo for the Mid-Atlantic and New England Fishery Management Councils. https://www.mafmc.org/s/Tab01_2020-SOE-Report_2020-04.pdf (2020; pp 33-49 of combined pdf), https://www.mafmc.org/s/Tab06_SOE-Mid-Atlantic-Risk-Assessment__2021-04.pdf (2021; pp 44-60 of combined pdf)

October 2017-2021: Lead author, Mid-Atlantic Fishery Management Council Ecosystem Approach to Fisheries Management Risk Assessments. https://www.mafmc.org/s/Tab08_EAFM-Risk-Assessment.pdf (2017),  https://www.mafmc.org/s/MAB_RiskAssess_08_18.pdf (2018), https://www.mafmc.org/s/Tab10_EAFM_Updates_2019-04.pdf (2019), https://www.mafmc.org/s/Tab01_2020-SOE-Report_2020-04.pdf (2020; pp 50-62 of combined pdf), https://www.mafmc.org/s/Tab06_SOE-Mid-Atlantic-Risk-Assessment__2021-04.pdf (2021; pp 61-81 of combined pdf)

December 2019: Mid-Atlantic Fishery Management Council Ecosystem Approach to Fisheries Management Conceptual models and visualizations. https://nefsc.github.io/READ-SSB-DePiper_Summer_Flounder_Conceptual_Models/sfconsmod_riskfactors_subplots.html https://nefsc.github.io/READ-SSB-DePiper_Summer_Flounder_Conceptual_Models/sfconsmod_final_2col.html
 
April 2016: Contributor to Mid-Atlantic Fishery Management Council Ecosystem Approach to Fisheries Management Guidance Document (1st Draft). http://www.mafmc.org/s/Tab05_EAFM.pdf 

February 2016: Lead author for A Framework for Incorporating Species, Fleet, Habitat, and Climate Interactions into Fishery Management: A DRAFT White Paper to Inform the Mid-Atlantic Fishery Management Council. http://www.mafmc.org/s/02_DRAFTInteractionsWhitePaper_bbdraft.pdf 

April 2015: Contributor/Editor for Climate Change and Variability: A White Paper to Inform the Mid-Atlantic Fishery Management Council on the Impact of Climate Change on Fishery Science and Management. http://www.mafmc.org/s/Climate-Change-and-VariabilityWhite-Paper_BBook_second_draft_final.pdf 

December 2014: Contributor to Managing Forage Fishes in the Mid-Atlantic Region: A White Paper to Inform the Mid-Atlantic Fishery Management Council. http://www.mafmc.org/s/3_ForageWhitePaper_Nov2014_draft3-2.pdf 

November 2010: Lead author, Bering Sea Ecosystem Assessment, and Ecosystem Considerations SAFE co-editor: http://access.afsc.noaa.gov/reem/ecoweb/Eco2010.pdf 

October 2010: GOA Crab Protection Closures Environmental Assessment, contributed ecosystem effects analysis in section 4: http://www.fakr.noaa.gov/npfmc/current_issues/bycatch/GOAcrab.pdf 

September 2010: Ecosystem Considerations SAFE editor (while S. Zador was on leave): ftp://ftp.afsc.noaa.gov/afsc/public/Plan_Team/ecosystem.pdf

November 2009: Ecosystem Assessment coauthor: http://www.afsc.noaa.gov/refm/docs/2009/ecosystem.pdf  

September 2008: Arctic Fishery Management Plan Environmental Analysis; contributed Arctic Ecosystem Description, Chapter 8, Section 1: http://www.fakr.noaa.gov/analyses/arctic/ArcticFMP_EA1108.pdf 

December 2007: Aleutian Islands Fishery Ecosystem Plan; Ecosystem Plan Team member, contributed food web analyses: http://www.fakr.noaa.gov/npfmc/current_issues/ecosystem/AIFEPbrochure1207.pdf 

1999-2006: Stock assessment reports (see above under Professional History) contributed to North Pacific Fishery Management Council Stock Assessment and Fishery Evaluation Reports:  https://archive.fisheries.noaa.gov/afsc/REFM/stocks/Historic_Assess.htm
 
# Presentations 

Speaker, moderator, and symposium organizer at regional to international professional meetings (American Fisheries Society, ASLO, Conservation Biology, Ecological Society of America, GLOBEC, IMCC, World Fisheries Congress, ICES Annual Science Conference) 1995-present. Invited speaker for public lecture series, teacher workshops, and university courses and departmental seminars.

## Talks presented at professional meetings

*Selected presentations, past 15 years*

Invited:

* 2021 

American Fisheries Society (AFS) Annual meeting, November 2021
    - connecting the dots climate informed management symposium
    - ecological forecasting symposium
    

    
* 2020 National SSC invited keynote speaker, meeting postponed to 2022

* 2019
SK Gaichas, GK DePiper, R Seagraves, B Muffley, S Lucey. There is no I in EAFM: adapting Integrated Ecosystem Assessment for Mid-Atlantic Fisheries Management. 2019 Coastal and Estuarine Research Foundation Biennial Conference. (Presented by Sean Lucey due to conflict with review panel)

* 2018
Gaichas et al herring control rules ICES ASC •	ICES ASC, September 2018 “Are any herring harvest control rules good for both fisheries and predators?”

* 2017

Sarah K. Gaichas, Richard Seagraves, Jessica Coakley, Geret DePiper, Vincent G. Guida, Jonathan Hare, Paul Rago and Michael Wilberg. The Mid-Atlantic Fishery Management Council’s Framework for Incorporating Species, Fleet, Habitat, and Climate Interactions into Fishery Management: Initial Risk Assessment. 2017 AFS Annual Meeting, Tampa FL. Coauthor on 5 other presentations at this meeting, and presented one for a coauthor.

coauthor on ICES presentations

IMBER IMBIZO stakeholders MSE lessons •	IMBER IMBIZO, October 2017 (“Getting on the same page, or at least in the same library: lessons in communication from a stakeholder-driven MSE for Northeast US Atlantic herring”)


* 2016
REDUS kickoff?

* 2015

•	NSAW, August 2015 (“Combining stock, multispecies, and ecosystem level status determination criteria: a worked example”); coauthor on ICES ASC abstract for Sept 2015 that used this material. 
•	AFS, August 2015 (“Multispecies stock assessment for Georges Bank: model development, performance testing, and multimodel inference”); Sarah K. Gaichas, Michael Fogarty, Robert J. Gamble, Sean M. Lucey, Laurel Smith, Charles Perretti, Gavin Fay coauthor on ICES ASC poster using this info. Coauthor on 2 other AFS abstracts.

AFS Portland

NSAW Portland
NOS NOAA modeling uncertainty workshop

•	ICES WGSAM, October 2014 (“Mixed fisheries in multispecies and ecosystem models” presented remotely)
•	Chesapeake Bay Forage Workshop, Nov 2014 (“Ecosystem metrics for fisheries management: 
examples and suggestions”)
•	ICES WKRISCO, Nov2014 (“WGNARS: IEA progress, opportunities, and challenges”)
•	NMFS-OAR Uncertainty Workshop, April 2015 (“Uncertainty in Ecosystem Models: Taxonomy and Treatment”)
•	NOAA GFDL Application of Seasonal to Decadal Climate Predictions for Marine Resource Management Workshop, June 2014 (“Incorporating Climate into Ecosystem Based Fisheries Management”)


* 2014
ICES ASC

•	ICES Symposium Ecological Basis of Risk Analysis for Marine Ecosystems, June 2014 (Northeast US fish community climate risk assessment—presented by Wes Patrick)
•	Ecological Forecasting Web-workshop, July 18 (Forecast use in multispecies assessments and ecosystem models, informal presentation with Rob Gamble and Tim Miller)
•	AFS, Canada, August 2014 (coauthor on Sean Lucey's talk, not attending the meeting)
•	ICES ASC, Spain, September 2014 (Ecosystem indicator analysis for risk assessment and IEA)


* 2013
•	ICES Gadoid Symposium, October 2013 (talk prepared, canceled due to govt shutdown)
•	ASLO Aquatic Sciences Meeting, New Orleans, LA, February 2013 (CAMEO Production modeling synthesis paper—presented by Tom Miller due to our travel restrictions) 
•	Data Visualization Workshop, Portland ME, February 2013 (Visualizing complexity; invited)
•	ICES WCSAM, Boston, MA, July 2013 (prepared abstract describing current multi-model, multispecies assessment project for Georges Bank, poster created and presented by Rob Gamble). 
•	Informal: Co-presented modeling work at EcoAP Research Retreat, December 2012 and drafted modeling report section with Rob Gamble


* 2012

•	World Fisheries Congress, May (CAMEO Production modeling synthesis paper)
•	AIFRB Symposium, June (Alaska “beer belly” paper)
•	ESA meeting, August (Multispecies size structured spatial assessment model development)
•	Supporting author on oral presentations at AFS (2 talks) and ICES (1 talk)


* 2011

-Presented three talks at the Lowell Wakefield Ecosystems Symposium in
Anchorage, AK, November 8-11, 2010: "From the Aleutians to the Arctic: Integrating Ecosystem Approaches within the Alaska Fishery Management Process" (for Stephani Zador), "Assembly Rules for Aggregate-Species Production Models: Simulations In Support of Management Strategy Evaluation" (with Mike Fogarty et al.), and "What Drives Dynamics in the Gulf of Alaska? Integrating Hypotheses of Species, Fishing, and Climate Relationships Using Ecosystem Modeling"

Invited keynote in PICES FUTURE workshop on Ecosystem Indicators, April 2011 in Honolulu, HI. 
AFS

* 2010

* 2009

* 2008

* 2007

*	Invited participant in Society for Conservation Biology symposium on ecosystem consequences of fishing in Port Elizabeth, South Africa (2007) and coauthor on a related review publication.

FY11:

-Presented "Marine ecosystems science and management from Alaska to Puget Sound" with Chris Harvey (NWFSC) at the Science Cafe, an informal public lecture series in Seattle sponsored by Pacific Science Center, December 7, 2010
-Presented Central GOA dynamic model simulations of whale restoration, jellyfish fluctuations, and no fishing at GLOBEC Pan-Regional Synthesis PI meeting, Feb 28-March 4, 2011
-Presented "Diet data products for stock assessment" and co-organized the Stock-specific ecosystem indicators workshop hosted by AFSC, April 7, 2011
-Invited to present "Uncertainty in Ecosystem Indicators: known knowns, known unknowns, and unknown unknowns" at the PICES FUTURE workshop on ecosystem indicators in Honolulu, HI, April 26-28, 2011

FY10:
-Invited presenter at UW SAFS “Young Investigator” departmental seminar on Dec 3, 2009: "Where does ecosystem modeling fit into fisheries management? Examples from the North Pacific”
- Co-led UW-NMFS Fisheries Think Tank on Dec 8, 2009: “Combining single species and ecosystem modeling results in ecosystem assessments” with Kerim Aydin and Isaac Kaplan (NWFSC)
- Co-presented with Kerim Aydin and Steph Zador a review of SSC comments and proposed analyses and processes to move forward NPFMC SSC Workshop on Ecological and Economic Indicators, February Council Meeting
- Participated in CAMEO Predator-Prey Interactions workshop in Corvallis, OR, March 16-18, 2010: presented three slide “lightning talk” on GOA food web modeling and management applications. - Presented “Ecosystem Considerations in the North Pacific: Groundfish SAFE &Aleutian Islands FEP” at the Ecosystem Science and Management workshop in Silver Spring, MD March 24-25, 2010.
- Presented “Fishing policy simulations in end-to-end models” at CAMEO end to end modeling workshop, Woods Hole MA, April 18-22 2010. Also, summarized discussions for two breakout groups and contributed meeting notes. - Presented GOA ecosystem overview and data summary, and examples of fitting our models to time series data at CAMEO stock production modeling workshop, Woods Hole MA, May 10-14

- Presented “Arctic ecosystem knowledge, where do we stand and where do we go from here?” in the AFSC seminar series, 19 May 2010. - Organized Alaska group analyses and presented work completed to date for ecosystem comparisons for the GLOBEC Pan-Regional Synthesis PI meeting in Seattle, May 25-27, 2010 - Presented “From the Aleutians to the Arctic: integrating ecosystem approaches within the Alaska fishery management process” at the AFS meeting in Pittsburgh, PA, 12-16 September within the Ecosystem Approach to Fishery Management symposium organized by Mike Fogarty et al. (NEFSC).
- Presented (for K. Aydin) “The use of models in ecosystem assessment processes in Alaska (experiences from the Alaskan Fisheries Council Process” at the AFS meeting in Pittsburgh, PA, 12-16 September within the Ecosystem Approach to Fishery Management symposium organized by Mike Fogarty et al. (NEFSC).
- Presented Ecosystem Considerations SAFE to the joint Groundfish Plan Teams, 22 September 2010.
FY09:
- Presented “Complex systems models of trophic and fishing interactions with implications for research and management” at GLOBEC Open Science Meeting in Victoria BC, June 2009
- Presented “Gulf of Alaska food web modeling and predation mortality estimates: information for single-species M” at the national workshop on Natural Mortality, August 2009
- Presented analysis of comparisons of single species and ecosystem model results across NMFS Science Centers at NEMoW II meeting in Annapolis, August 2009
- Presented work completed to date for ecosystem comparisons for the GLOBEC Pan-Regional Synthesis PI meeting in Seattle, September 2009 

FY08:
- Invited expert at Asia-Pacific Economic Cooperation (APEC) workshop on developing a network on deep sea resources and fisheries, October 3-5 2007 in Lima, Peru; presented talk on Alaskan fisheries and deep-sea resources including slope surveys in all three ecosystems.
- Presented results of ecosystem threshold modeling at UW mini-workshop October 24, 2007
- Presented ecosystem modeling talk at Pacific Science Center Teacher workshop October 26, 2007
- PICES Annual meeting, Victoria, B.C., October 30, 2007; presented my own paper on ecosystem level thresholds as well as Ivonne Ortiz's paper on Aleutian Islands spatial food web modeling. - UW Career Week panelist for marine science careers January 29, 2008
- Science on Tap May 6, 2008; informal evening outreach presentation on ecosystem modeling and fishery management, sponsored by Pacific Science Center
- Marine Science Weekend at Pacific Science Center May 15-18, 2008; collaborated to develop and present a food-web/ecology related demonstration (did first presentation, May 15)
- Presented ecosystem modeling results to Ray Hilborn's UW class, May 19, 2008
- Advances in Marine Ecosystem Modelling Research (AMEMR) symposium in Plymouth, UK June 22-26, 2008; presented "Identifying critical interactions and thresholds in eastern north Pacific ecosystems through model simulations".

FY07:
-PICES meeting Yokohama Japan October 13-19, 2006. Presented ecosystem network comparison methods for EBS, GOA, and California Current during CFAME workshop.
-Society for Conservation Biology meeting invited symposium speaker July 1-5, 2007 in South Africa. -AFS: presented a talk and a poster (for Grant Thompson) in the symposium “Ecosystem modeling applications in fishery management” for September 2-6, 2007 meeting in San Francisco -MENU workshop, Bergen Norway, March 6-9 presented GOA ecosystem description, demonstrated comparison methods used in Alaska.
-Co-leader of UW Mini-workshop on fitting ecosystem models to data, February 20 2007.
-NEMoW workshop, Santa Cruz CA August 29-31, 2007, gave presentation on Ecosim and contributed to presentation on Ecopath, co-leader of breakout group 1, and rapporteur. -Presented research on computation involved in ecosystem modeling to a group of visiting junior college teachers from Fresno City College and other schools touring the NOAA facility August 9, 2007. -Presenter at AFSC fish booth at the Seattle Fall Fisherman’s Festival, September 22, 2007.


## Posters presented at professional meetings

•	Poster presentations, National SSC meeting, January 2018 ("Which MSE should you do first? Prioritizing MSE investments using risk assessment under an Ecosystem Approach to Fishery Management", and “Getting on the same page, or at least in the same library: lessons in communication from a stakeholder-driven management strategy evaluation for northeast US Atlantic herring”)

## Other seminars/presentations

Fishery Management Councils: SOE

Fishery Management Councils Management stategy evaluation stakeholder workshops

Marine Resource Education Program

Maine Fishermen's Forum

OneNOAA EBFM series

AFSC Groundfish series

NEFSC seminar series

UMass seminar series

IMR/NOAA bilateral meetings

IMR departmental and institute seminars

Invited
* Keynote on multispecies modeling best practices for Multispecies Modeling Applications Workshop, June 2021
* IEA and SOE for MACO Ocean forum May 2021
* Using ecosystem info in management for DFO Ecosystem WG February 2021
* ICES WKRPchange US reference points and herring HCR (2 presentations) Sept 2020
* MSE and DMDU for Ecosystem WG NOAA SAB July 2020
* Fisheries Forum April 2017

[US Reference Points Introduction](https://noaa-edab.github.io/presentations/20200921_USrefpts_WKRPchange_Gaichas.html#1), September 21,  2020, ICES Workshop on Reference Point Changes

[Herring HCR accounting for its role as forage](https://noaa-edab.github.io/presentations/20200923_herringWKRPchange_Gaichas.html#1), September 23, 2020, ICES Workshop on Reference Point Changes

[Species Interactions in Multispecies models](https://noaa-edab.github.io/presentations/20201008_FIT_MSmods_Gaichas.html#1), October 8 2020, NOAA Fisheries Integrated Toolbox Technical meeting 

[Multispecies models: US update](https://noaa-edab.github.io/presentations/20201012_WGSAM_USupdate_Gaichas.html#1), October 12, 2020, ICES Working Group on Multispecies Assessments

[Integrating Climate into SOE and Model Skill Assessment](https://docs.google.com/presentation/d/184Af-afIIv-aGSRT6-Ed9SRmqf5n8YOO/edit#slide=id.p1), November 17, 2020, Northeast Fisheries-Climate PI workshop

“State of the Ecosystem 2020” for ICES WGNARS, May 8 (pre-recorded in April): https://noaa-edab.github.io/pre sentations/20200508_US_SOEUpdate_Gaichas.html

"US Council update (EAFM/EBFM): for ICES WGNARS, May 8 (pre-recorded in April): https://noaa-edab.github. io/presentations/20200508_USCouncilUpdate_Gaichas.htm

“Climate and Fisheries” for National Academy of Sciences Ocean Studies Board panel, May 22, 2020: https://noaa- edab.github.io/presentations/20200522_NASOSB_Gaichas.html

“Deep Uncertainty and MSE” for NOAA Science Advisory Board’s Ecosystem Sciences and Management Working Group, July 16, 2020: https://noaa-edab.github.io/presentations/20200716_MSEUncertainty_Gaichas.html

* 2018

•	ICES WGNARS, March 2018 (“US Risk Assessment and Ecosystem Reporting Update”)

•	International TAC workshop in Xiamen China, April 2018:“Can we integrate ecosystem and multispecies interactions into TACs and other fishery management?”


*  2017

•	ICES WGNARS, March 2017 (“US IEA Update”)
•	ICES WGSAM, October 2017 (“Update on US modeling activities”, “Indicator based risk assessment”, “Herring MSE update”, “Linking mass balance models” for Sean Lucey, “GB multispecies management procedure” for Mike Fogarty, “Model structural comparisons” for Robert Wildermuth, “ICES update” for Amanda Hart, “Mulitple models for sardine” for Isaac Kaplan, “ACLIM project” for Kirstin Holsman)
•	NSF Wind Energy Workshop, November 2017 (“IEAs”)
•	QUEST webinar, December 2017 (“The Dream and the Reality: Meeting decision-making times frames while incorporating ecosystem and economic models into management strategy evaluation”)

*  2016

•	ICES WGSAM, October 2016 (“Update on US MSE activities and Atlantis Poseidon Adventure project”, “The state of ecosystem indicator reference points for fishery management” for Jamie Tam, “Tradeoffs between commercial fishing fleets, marine mammals and finfish biomass” for Laurel Smith, and “Mid-Atlantic Ecosystem Approach to Fisheries Management”)

•	Council presentations on herring MSE and EAFM reported under elements 3 and 4


# Funded research

Internal grants (multispecies modeling, MSE support, fish energy content indicator)
International fellowship 
COCA

* Co-PI on modeling portion of the North Pacific Research Board’s GOA Integrated Ecosystem Research Plan.
*	Investigator with 22 others on the interdisciplinary "Collaborative Research: GLOBEC Pan-Regional Synthesis: End-to-end Energy Budgets in US-GLOBEC Regions" project comparing four marine ecosystems in the Southern Ocean, in the Pacific off Alaska and the U.S. West Coast, and in the Atlantic off the U.S East Coast (2008-2011).

# Professional Affiliations/Activities/Service

## Professional societies

Current member: American Fisheries Society ([AFS](https://fisheries.org/)), American Institute of Fishery Research Biologists ([AIFRB](https://www.aifrb.org/))

Past member: Ecological Society of America, The Society for Conservation Biology, American Society of Ichthyologists and Herpetologists

## Editorial positions

Guest editor, CJFAS special issue MSE 2018-2019

Guest editor, MEPS special issue comparative system productivity 2011-2012

## Membership on panels

Chaired multispecies model key run reviews for ICES WGSAM, October 2020 (North Sea), October 2019 (Baltic Sea). Participant in WGSAM model key run reviews 2015-2018.

Invited review panelist:

*  Atlantic States Marine Fisheries Commission Atlantic Menhaden stock assessment and ecological reference points, November 2019.

*  South African Assessment review, November 2018. MSE-based management procedures for 4 species were reviewed over the course of a week by an international panel. (Also invited for 2019 panel, but declined due to conflicts).

*  International Whaling Commission, JARPN II February 2016 (Tokyo, Japan), NEWREP-NP February 2017 (by correspondence)

*  Two NOAA NMFS Ecosystem Science Program reviews: Alaska (May 2016) and Northwest (July 2016)

* 2015-2013?

*  Invited external reviewer for the Canada Maritimes Region Ecosystem Research Initiative (March 2012)


## Advisory services

    Active
* Mid-Atlantic FMC Scientific and Statistical Committee, (2014-present); determining Allowable Biological Catch (ABC) advice for fishery management and other scientific support. 
* Member and past co-chair of the ICES Working Group on North Atlantic Regional Seas; developing ecosystem indicators and integrated ecosystem assessments for Canadian and U.S. waters.
* Co-Chair of the ICES Working Group on Multispecies Assessment Methods: reviewing and contributing to multispecies and ecosystem modeling progress across ICES regions. 
* Advisory partner in the European Commission's PANDORA project (2018-2022), which is designing toolsets for new dynamic ocean resource assessments and exploitation.


    Past
All the Council stuff

Herring assesment WG 2018

Invited speaker/participant, Multispecies Fisheries Management Workshop, Fujian, China 2018

ICES WKs
WKRPchange 2020
WKRISCO 2014

* Advisory partner in the European Commission's MYFISH project (2012-2015), which defined alternative approaches to Maximum Sustainable Yield including ecological, social, and economic dimensions.


CB Forage Fish 2014

* Member of the Aleutian Islands Fishery Ecosystem Plan Team (2007-2011) and the Gulf of Alaska Fishery Management Plan Team (2000-2011) for North Pacific Fishery Management Council. Co-authored the Aleutian Islands Fishery Ecosystem Plan (2007) and regularly co-authored the Stock Assessment and Fishery Evaluation introduction and summary for the Gulf of Alaska. Authored Ecosystem Description and contributed analyses for management alternatives for the Arctic Fishery Management Plan (2009).
* Co-organizer of the Bering Sea Ecosystem Synthesis Team (2010), a multi-disciplinary group charged with identifying important ecosystem indicators for fishery management and writing an eastern Bering Sea ecosystem assessment for the North Pacific Fishery Management Council. 
*	Invited participant in Asia-Pacific Economic Cooperation International workshop on developing a network for deep sea research in Lima, Peru (2007). Presented information on North Pacific ecosystems, fisheries, and management, and helped develop a preliminary framework for an information-sharing network. 



## Workshops organized/chaired; sessions organized/chaired at professional meetings

    In Progress

ICES/PICES 2022 Small Pelagics Symposium Lisbon, Portugal; S6 Reconciling Ecological Roles and Harvest Goals: Development and Testing Management Strategies to Enhance Marine Ecosystem Services. https://meetings.pices.int/meetings/international/2022/pelagic/program#S6 Convenors: Sarah Gaichas, Cecilie Hansen, Isaac Kaplan, Richard Nash

World Fisheries Congress 2021 (Adelaide, Australia/Virtual); The land of plenty: Advances and future directions in population dynamics modeling to support fishery management. Six consecutive, non-overlapping sub-sessions are planned, including (conveners in parentheses): general stock assessment methodology and statistics (André Punt, James Thorson); spatial models (Daniel Goethel, Aaron Berger); ecosystem dynamics and climate change (Éva Plagányi, Caleb Gardner, Sarah Gaichas); data-limited methods (Kristen Omori, Jason Cope, Natalie Dowling); management strategy evaluation and biological reference points (Carryn de Moor, Richard McGarvey, Ann Preece); and a summary discussion (Mark Maunder, Patrick Lynch).

    Past

National Ecosystem Modeling Workshop (NEMoW) 2019 St Petersburg; Steering committee member and workshop planner
– Developed NEFSC overview presentation (presented by Scott Large)
– Developed atlantisom tool demonstration (presented by Christine Stawitz)

American Fisheries Society annual meeting (AFS) 2018 Atlantic City; Emergent Properties of Aquatic and Marine Ecosystems Due to Climate Change: An Overview of the Current State of Knowledge. Organizers: Michelle D. Staudinger Abigail J. Lynch, Sarah K. Gaichas

AFS 2017 Tampa; Closing the Loop: Stakeholder Involvement in the Management Strategy Evaluation (MSE) Process—Advancing Management Strategy Evaluation. Organizers: Sean M. Lucey, Sarah K. Gaichas, John F. Walter III, Daniel Goethel, Aaron Berger, and Patrick Lynch 

AFS 2017 Tampa; Under Pressure: Defining Harvest Strategies that Account for Biological, Environmental or Anthropogenic Spatiotemporal Complexity—Advancing Management Strategy Evaluation. Organizers: Daniel Goethel, Aaron Berger, Patrick Lynch, Sarah K. Gaichas, John F. Walter III, and Sean M. Lucey 

AFS 2015 Portland, OR; Multispecies models (including humans!)—where have we been and where do we need to go? Organizers: Sarah K. Gaichas, Kirstin Holsman, Alan Haynie and Geret DePiper

AFS 2011 ; What Influences Fisheries Production? Comparing the Effects of Environmental, Fishing, and Food Web Forcing Across Large Marine Ecosystems

Comparative Analysis of Marine Ecosystem Organization (CAMEO) Stock Production Modeling Workshop II in Woods Hole, 2011. 

Society for Conservation Biology's International Marine Conservation Conference (IMCC) 2009 Victoria, BC;	Bridging the gap between fisheries and marine conservation to advance ecosystem-based management 

AFS 2007 San Francisco; Ecosystem modeling applications in fishery management

## Teaching and mentoring

    Graduate-level

**Adjunct professor**  *University of Massachusetts, Dartmouth* 

One current College of William and Mary (Virginia Institute of Marine Science) PhD student committee  
One current UNH PhD student committee

2021  (Exam) PhD student committee at the University of Massachusetts, Dartmouth  
2020  External examiner for a PhD student at University of Iceland  
2019  (Exam) PhD student committee at the University of Massachusetts, Dartmouth   
2019  External examiner for a PhD student at Dalhousie University  
2016	External examiner for a PhD student at the University of Cape Town   
2016	(Exam) PhD student committee at the University of Washington   
2015	(Exam) PhD student committee at Oregon State University  


    Undergraduate-level

IN FISH program 2021 Internship mentor

Shoals Marine Lab summer course and lab guest lecturer 2021, 2019

WPI REU program (twice), organized lab visits and provided guidance, background, and data for student projects

Invited course and lab guest lecturer for Rutgers 2020 summer course, postponed due to COVID


*	Guest lecturer: Woods Hole Partnership in Education Program (PEP), June 2015; Worcester Polytechnic Institute Research Experience for Undergraduates in Applied Mathematics, June 2014; WHOI Fisheries Oceanography Course, April 2014, Rutgers Ecology and Evolution Dept. Seminar, March 2013, SMAST Fisheries Oceanography Dept. Seminar, February 2012.

*	Regular participant in Alaska Fisheries Science Center scientific education and outreach activities, including Pacific Science Center programs, Minorities in Marine Science Undergraduate Program presentations, elementary, secondary, and college educator workshops, internship application review and placement.  


# Agency Service

## Committees

National MSE WG Chair

MAFMC SSC (here or advisory?)

* Participated in National EBFM workshop, July 2020
    +  Contributed NEFSC info to ESR presentation
    +  Represented NEFSC in ESR and other discussions



## Other

# Other relevant information

March 2020 - present: considerable time was needed for dependent care (elementary age kids and elder care). I often used the excused absence for dependent care.

# References

Vera Agostini
FAO
Deputy Director
Fisheries and Aquaculture Policy and Resources Division
Food and Agriculture Organization of the United Nations
Viale delle Terme di Caracalla
00153 Rome, Italy.
Vera.Agostini@fao.org
P. +39 06570 50183 ; F. +39 06570 53020
Skype vera.agostini


Jane DiCosimo
Retired, formerly 
13411 Norden Dr, Silver Spring, MD 20906
Or if it would help more
34198 Eaglet Way Soldotna AK 99669



Daniel Howell
Institute of Marine Research
P.O. Box 1870 Nordnes
NO-5817 Bergen, Norway
daniel.howell@hi.no
+47 92063885


Jason Link


