---
address: |
  NOAA NMFS Northeast Fisheries Science Center
  166 Water Street, Woods Hole, MA 02543
date: "`r format(Sys.time(), '%B %Y')`"
email: Sarah.Gaichas@noaa.gov
github: sgaichas
name: Sarah
output:
  vitae::awesomecv:
    keep_tex: yes
  vitae::moderncv: default
phone: +1 508 495 2016
position: Research Fishery Biologist ZP-IV
surname: Gaichas
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(vitae)
```

# Educational History

```{r}
library(tibble)

#rorcid::orcid_auth()

orcid_data <- do.call("rbind",
    rorcid::orcid_educations("0000-0002-5788-3073")$`0000-0002-5788-3073`$`affiliation-group`$summaries
)

orcid_data %>%
  detailed_entries(
    what = `education-summary.role-title`,
    when = glue::glue("{`education-summary.start-date.year.value`} - {`education-summary.end-date.year.value`}"),
    with = `education-summary.organization.name`,
    where = glue::glue("{`education-summary.organization.address.city`},{`education-summary.organization.address.region`}"),
  )

# tribble(
#   ~ Degree, ~ Year, ~ Institution, ~ Where,
#   "Informal studies", "1889-91", "Flying University", "Warsaw, Poland",
#   "Master of Physics", "1893", "Sorbonne Université", "Paris, France",
#   "Master of Mathematics", "1894", "Sorbonne Université", "Paris, France"
# ) %>% 
#   detailed_entries(Degree, Year, Institution, Where)

```


# Professional History

**Research Fishery Biologist**                   		*NOAA/NMFS Northeast Fisheries Sci. Center, Woods Hole, MA*

	Ecosystem Assessment, MSE		September 2011 - present
Contribute to integrated ecosystem assessments and management strategy evaluations (MSEs) for the Northeast U.S. continental shelf by analyzing research survey, fishery, oceanographic, and economic data, and by developing and testing multispecies assessment models and ecosystem indicators. Conduct research on ecosystem based management reference points balancing multiple objectives. Chair National and Northeast MSE working groups to build capacity and coordinate work plans. Lead editor, State of the Ecosystem reports. Provide results in peer-reviewed literature and to the New England and Mid-Atlantic Fishery Management Councils. 

**Research Fishery Biologist**                   		*NOAA/NMFS Alaska Fisheries Science Center, Seattle, WA*

	Ecosystem Modeling			June 2006 – September 2011
Established an ecosystem context for fishery management through analysis of food web, fishery, and research survey data combined with ecosystem modeling.  Evaluated the cumulative effects of fishing, climate change, and other factors on the marine ecosystem and species within it. Conducted independent research on ecosystem level indicators and thresholds and provided results in peer-reviewed literature and to the North Pacific Fishery Management Council. Participated in research surveys at sea. 

	Stock Assessment			November 1998 – June 2006
Assessed the condition of commercially exploited fish stocks through analysis of fishery and research survey data combined with population dynamics modeling. Evaluated the effects of fishing on non-target species and the marine ecosystem.  Conducted Ph.D. research. Attended 2001 Complex Systems Summer School, Santa Fe Institute, Santa Fe NM. Participated in research surveys at sea.

Stock assessments:  1999 Gulf of Alaska (GOA) and 2000 – 2004 Bering Sea Aleutian Islands (BSAI) “Other species” (all species of skates, sharks, sculpins, octopi, and squids), 2000-2005 BSAI squids, 2005 BSAI skates, 2003-2005 GOA skates, 2000-2005 GOA Thornyheads, and 2006 GOA squids.  


	Observer Program			March 1997 – October 1999
Conducted statistical analyses of fishery observer data, evaluated present and proposed data collection methods through directed research, and examined uses of observer data in multiple applications. 

# Awards

* End-year performance bonuses each year, 2006-2020.
* Individual cash award for State of the Ecosystem reporting, August 2020.
* Individual cash award for Acting Branch Chief performance, July 2017.
* ICES/MYFISH symposium Targets and Limits for Long Term Fisheries Management Most Inspiring Paper, October 2015. 
* Department of Commerce “Cash-In-A-Flash” award, December 2011.
* Department of Commerce Special Act Award: Alaska Fisheries Science Center Employee of the Year, 2010.
* Department of Commerce Special Act Award for service on North Pacific Council Gulf of Alaska Plan Team and for ecosystem research contributions, July 2005.
* Department of Commerce Special Act Award for innovations in nontarget species management and ecosystem modeling, 2004.
* National Association of Environmental Professionals Award for Environmental Excellence in NEPA for assistance with the Steller Seal Lion Protection Measures SEIS, June 2003.
* Mote International Symposium on Confronting Trade-offs in the Ecosystem Approach to Fisheries Management Young Investigator Award, November 2002.
* Department of Commerce Special Act Award for service as the Nontarget and Prohibited Species Team Leader in the development of the Draft Programmatic SEIS for groundfish fisheries in Alaska, September 2000. 
* Department of Commerce “Cash-In-A-Flash” awards, March and June 1998. 
* College of William and Mary William J. Hargis Jr. Fellowship Award and Dean's Fellowship. April, 1995.

# Publications

## Papers published in peer-reviewed journals

```{r bib, warning=FALSE, results='asis'}

if(!file.exists("gaichas.bib")){
scholar::get_publications("C_gOAZ4AAAAJ&hl") %>% 
  #filter(year > 2014) %>% 
  arrange(desc(year)) %>%
  distinct(title, .keep_all = TRUE) %>%
  transmute(bibtype = "Article", author = as.character(author),
            title = as.character(title),
            journaltitle = as.character(journal), year, key = row_number()) %>% 
  RefManageR::as.BibEntry() %>% 
  RefManageR::WriteBib("gaichas.bib")
}

# this bib from zotero; google scholar brought in conference papers, etc
bib <- RefManageR::ReadBib("gaichas.bib")
#sort(bib, sorting = "ydnt")
RefManageR::BibOptions(sorting = "ydnt", 
                       bib.style = "authoryear", 
                       max.names = 50,
                       style = "markdown",
                       no.print.fields = c("abstract", "copyright", "issn", "url", "language","urldate","month","file", "shorttitle", "note", "keywords"))
RefManageR::NoCite(bib)

newline <- list(rep("  \n", length(bib)))
RefManageR::BibEntry()
#bib <- mapply(append, RefManageR::BibEntry(bibtype = "doi"), newline, SIMPLIFY = FALSE)
RefManageR::PrintBibliography(bib)

#bibliography_entries("gaichas.bib")%>%
#  select(author, year, title, journal, pages, url) %>%
#  arrange(desc(year, author$family))
```

## Papers submitted to peer-reviewed journals

Staudinger, M., Lynch, A.,  Gaichas, S., Fox, M., Gibson-Reinemer, D., Langan, J., Teffer, A., Thackeray, S., and Winfield, I.  Accepted April 2021. How does climate change affect emergent properties of aquatic ecosystems? Fisheries.

Kaplan, I.C., S.K. Gaichas, C.C. Stawitz, P.D. Lynch, K.N. Marshall, J.J. Deroba, M. Masi, J.K.T. Brodziak, K.Y. Aydin, K. Holsman, H. Townsend, D. Tommasi, J.A. Smith, S. Koenigstein, M. Weijerman and J. Link. Management Strategy Evaluation: Allowing the Light on the Hill to Illuminate More than One Species. Original Research, Front. Mar. Sci. - Marine Fisheries, Aquaculture and Living Resources. 

## Other papers (non peer-reviewed journals, book chapters, symposium volumes)

Gaichas, S.K., C. Reiss, and M. Koen-Alonso. 2014. Ecosystem Based Management in high latitude ecosystems. Chapter 10, p. 277-324 in The Sea, Volume 16: Marine Ecosystem Based Management. Harvard U. Press.

## Books

## Reports

Northeast Fisheries Science Center State of the Ecosystem Reports; lead editor for Mid-Atlantic 2017-2021, lead editor for New England 2017-2020, supporting editor for New England 2021. 

* 2021 State of the Ecosystem: Mid-Atlantic; https://apps-nefsc.fisheries.noaa.gov/rcb/publications/SOE-MAFMC-2021-508-Final.pdf
* 2021 State of the Ecosystem: New England; https://apps-nefsc.fisheries.noaa.gov/rcb/publications/SOE-MAFMC-2021-508-Final.pdf
* 2020 State of the Ecosystem: Mid-Atlantic; https://doi.org/10.25923/1f8j-d564
* 2020 State of the Ecosystem: New England; https://doi.org/10.25923/4tdk-eg57
* 2019 State of the Ecosystem: Mid-Atlantic; https://www.mafmc.org/s/Tab09_State-of-the-Ecosystem-Report_2019-04.pdf
* 2019 State of the Ecosystem: New England; https://s3.amazonaws.com/nefmc.org/11_SOE-NEFMC-2019.pdf
* 2018 State of the Ecosystem: Mid-Atlantic; https://www.mafmc.org/s/Tab04_State-of-the-Ecosystem-Apr2018.pdf
* 2018 State of the Ecosystem: New England; https://s3.amazonaws.com/nefmc.org/2_Ecosystem-Status-Report.pdf
* 2017 State of the Ecosystem: Mid-Atlantic; https://www.mafmc.org/s/Tab02_2017-04_State-of-the-Ecosystem-and-EAFM.pdf
* 2017 State of the Ecosystem: New England; https://s3.amazonaws.com/nefmc.org/2_2016-State-of-the-Ecosystem-Report.pdf

ICES. 2021. Working Group on Multispecies Assessment Methods (WGSAM; outputs from 2020 meeting). ICES Scientific Reports. 3:10. 231 pp. https://doi.org/10.17895/ices.pub.7695

ICES. 2021. Workshop of Fisheries Management Reference Points in a Changing Environment(WKRPChange, outputs from 2020 meeting). ICES Scientific Reports. 3:6. 39 pp.https://doi.org/10.17895/ices.pub.7660

ICES. 2019. Working Group on Multispecies Assessment Methods (WGSAM). ICES Scientific Reports.1:91. 320 pp. http://doi.org/10.17895/ices.pub.5758

ICES. 2018. Report of the Working Group on Multispecies Assessment Methods (WGSAM), 15–19 October 2018, Paris, France. ICES CM 2018/HAPISG:20. 89 pp.

ICES WGNARS

1-2 other NEMoW reports

Link, J.S., D. Mason, T. Lederhouse, S. Gaichas, T. Hartley, J. Ianelli, R. Methot, C. Stock, C. Stow, and H. Townsend. 2015. Report from the Joint OAR-NMFS Modeling Uncertainty Workshop. U.S. Dept. of Commer., NOAA. NOAA Technical Memorandum NMFS-F/SPO-153, 31 p.

Lucey, S. and S. Gaichas. 2015. Fish Communities. Section in Fogarty, M. (ed.), Ecosystem Status Report for the Northeast Large Marine Ecosystem http://www.nefsc.noaa.gov/ecosys/ecosystem-status-report/fish-communities.html 
Northeast Fisheries Science Center 2015. Figures for Ecosystem Status Report (Executive Summary). http://www.nefsc.noaa.gov/ecosys/ecosystem-status-report/ 

Link, J. S., T.F. Ihde, H. Townsend, K. Osgood, M. Schirripa, D. Kobayashi, S. Gaichas, J. Field, P. Levin, K. Aydin, G. Watters, and C. Harvey (editors). 2010. Report of the 2nd National Ecosystem Modeling Workshop (NEMoW II):  Bridging The Credibility Gap – Dealing With Uncertainty In Ecosystem Models. U.S. Dep. Commerce, NOAA Tech. Memo. NMFS-F/SPO-102, 72 p.

Conners, M.E., J. Cahalan, S. Gaichas, W.A. Karp, T. Loomis, and J. Watson. 2009. Sampling for estimation of catch composition in Bering Sea trawl fisheries. NOAA Tech. Memo. NMFS-AFSC-199, 42 p. 

Aydin, K., S. Gaichas, I. Ortiz, D. Kinzey, and N. Friday. 2007. A comparison of the Bering Sea, Gulf of Alaska, and Aleutian Islands large marine ecosystems through food web modeling. U.S. Dep. Commer., NOAA Tech. Memo. NMFS-AFSC-178, 298 p.

Gaichas, S.K., and J.C. Field, 2003. Comparative modeling approaches for exploited marine ecosystems in the North Pacific: what are these models good for? In Mace, P.M. (ed.), Proceedings of the seventh NMFS National Stock Assessment Workshop, (Re)building Sustainable Fisheries and Marine Ecosystems.  NOAA Tech. Memo NMFS-F/SPO-62.


## Others (Fishery Management Council Documents)

2020-2021: Lead author, State of the Ecosystem Request tracking memo for the Mid-Atlantic and New England Fishery Management Councils

2017-2021: Lead author, Mid-Atlantic Fishery Management Council Ecosystem Approach to Fisheries Management Risk assessments

December 2019: Mid-Atlantic Fishery Management Council Ecosystem Approach to Fisheries Management Conceptual model
 
April 2016: Contributor to Mid-Atlantic Fishery Management Council Ecosystem Approach to Fisheries Management Guidance Document (1st Draft). http://www.mafmc.org/s/Tab05_EAFM.pdf 

February 2016: Lead author for A Framework for Incorporating Species, Fleet, Habitat, and Climate Interactions into Fishery Management: A DRAFT White Paper to Inform the Mid-Atlantic Fishery Management Council. http://www.mafmc.org/s/02_DRAFTInteractionsWhitePaper_bbdraft.pdf 

April 2015: Contributor/Editor for Climate Change and Variability: A White Paper to Inform the Mid-Atlantic Fishery Management Council on the Impact of Climate Change on Fishery Science and Management. http://www.mafmc.org/s/Climate-Change-and-VariabilityWhite-Paper_BBook_second_draft_final.pdf 

December 2014: Contributor to Managing Forage Fishes in the Mid-Atlantic Region: A White Paper to Inform the Mid-Atlantic Fishery Management Council. http://www.mafmc.org/s/3_ForageWhitePaper_Nov2014_draft3-2.pdf 

November 2010: Lead author, Bering Sea Ecosystem Assessment, and Ecosystem Considerations SAFE co-editor: http://access.afsc.noaa.gov/reem/ecoweb/Eco2010.pdf 

October 2010: GOA Crab Protection Closures Environmental Assessment, contributed ecosystem effects analysis in section 4: http://www.fakr.noaa.gov/npfmc/current_issues/bycatch/GOAcrab.pdf 

September 2010: Ecosystem Considerations SAFE editor (while S. Zador was on leave): ftp://ftp.afsc.noaa.gov/afsc/public/Plan_Team/ecosystem.pdf

November 2009: Ecosystem Assessment coauthor: http://www.afsc.noaa.gov/refm/docs/2009/ecosystem.pdf  

September 2008: Arctic Fishery Management Plan Environmental Analysis; contributed Arctic Ecosystem Description, Chapter 8, Section 1: http://www.fakr.noaa.gov/analyses/arctic/ArcticFMP_EA1108.pdf 

December 2007: Aleutian Islands Fishery Ecosystem Plan; Ecosystem Plan Team member, contributed food web analyses: http://www.fakr.noaa.gov/npfmc/current_issues/ecosystem/AIFEPbrochure1207.pdf 

1999-2006: Stock assessment reports (see below) contributed to North Pacific Fishery Management Council Stock Assessment and Fishery Evaluation Report https://archive.fisheries.noaa.gov/afsc/REFM/stocks/Historic_Assess.htm
 
# Presentations 

Speaker, moderator, and symposium organizer at regional to international professional meetings (AFS, ASLO, Conservation Biology, ESA, GLOBEC, IMCC, World Fisheries Congress) 1995-present. Invited speaker for public lecture series, teacher workshops, and university courses and departmental seminars.

## Talks presented at professional meetings

Invited: 
* 2021 AFS 
    - connecting the dots climate informed management symposium
    - ecological forecasting symposium

OMG too many

## Posters presented at professional meetings

Fewer

## Other seminars/presentations

Fishery Management Councils: SOE

Fishery Management Councils Management stategy evaluation stakeholder workshops

Marine Resource Education Program

Maine Fishermen's Forum

OneNOAA EBFM series

AFSC Groundfish series

NEFSC seminar series

UMass seminar series

IMR/NOAA bilateral meetings

Invited
* Using ecosystem info in management for DFO Ecosystem WG
* MSE and DMDU for Ecosystem WG NOAA SAB





# Funded research

Internal grants
International fellowship
COCA

# Professional Affiliations/Activities/Service

* Mid-Atlantic FMC Scientific and Statistical Committee, (2014-present); determining Allowable Biological Catch (ABC) advice for fishery management and other scientific support. 
* Member and past co-chair of the ICES Working Group on North Atlantic Regional Seas; developing ecosystem indicators and integrated ecosystem assessments for Canadian and U.S. waters.
* Co-Chair of the ICES Working Group on Multispecies Assessment Methods: reviewing and contributing to multispecies and ecosystem modeling progress across ICES regions. 
* Advisory partner in the European Commission's PANDORA project (2018-2022), which is designing toolsets for new dynamic ocean resource assessments and exploitation.
* Advisory partner in the European Commission's MYFISH project (2012-2015), which defined alternative approaches to Maximum Sustainable Yield including ecological, social, and economic dimensions.
*	External reviewer for the Canada Maritimes Region Ecosystem Research Initiative (March 2012).
*	Guest lecturer: Woods Hole Partnership in Education Program (PEP), June 2015; Worcester Polytechnic Institute Research Experience for Undergraduates in Applied Mathematics, June 2014; WHOI Fisheries Oceanography Course, April 2014, Rutgers Ecology and Evolution Dept. Seminar, March 2013, SMAST Fisheries Oceanography Dept. Seminar, February 2012.
* Member of the Aleutian Islands Fishery Ecosystem Plan Team (2007-2011) and the Gulf of Alaska Fishery Management Plan Team (2000-2011) for North Pacific Fishery Management Council. Co-authored the Aleutian Islands Fishery Ecosystem Plan (2007) and regularly co-authored the Stock Assessment and Fishery Evaluation introduction and summary for the Gulf of Alaska. Authored Ecosystem Description and contributed analyses for management alternatives for the Arctic Fishery Management Plan (2009).
* Co-organizer of the Bering Sea Ecosystem Synthesis Team (2010), a multi-disciplinary group charged with identifying important ecosystem indicators for fishery management and writing an eastern Bering Sea ecosystem assessment for the North Pacific Fishery Management Council. 
*	Invited participant in PICES FUTURE workshop on Ecosystem Indicators, April 2011 in Honolulu, HI. 
*	Co-organizer of the Comparative Analysis of Marine Ecosystem Organization (CAMEO) Stock Production Modeling Workshop II in Woods Hole, MA, May 2011. Presented results at the International Marine Conservation Congress in Victoria BC (May 15, 2011) at the Ecosystem Studies of Subarctic Seas Open Science Meeting (May 24, 2011), and at the World Fisheries Congress in Edinburgh, Scotland (May 2012). 
* Participant in three CAMEO workshops in 2010: Role of predator-prey interactions in marine ecosystem organization in Corvallis, OR, March 2010; End to end modeling workshop in Woods Hole, MA, April 2010; and Stock production model fitting workshop in Woods Hole MA, May 2010. 
* Co-PI on modeling portion of the North Pacific Research Board’s GOA Integrated Ecosystem Research Plan.
*	Investigator with 22 others on the interdisciplinary "Collaborative Research: GLOBEC Pan-Regional Synthesis: End-to-end Energy Budgets in US-GLOBEC Regions" project comparing four marine ecosystems in the Southern Ocean, in the Pacific off Alaska and the U.S. West Coast, and in the Atlantic off the U.S East Coast (2008-2011).
*	Invited participant in Asia-Pacific Economic Cooperation International workshop on developing a network for deep sea research in Lima, Peru (2007). Presented information on North Pacific ecosystems, fisheries, and management, and helped develop a preliminary framework for an information-sharing network. 
*	Invited participant in Society for Conservation Biology symposium on ecosystem consequences of fishing in Port Elizabeth, South Africa (2007) and coauthor on a related review publication.  
*	Participant in the Marine Ecosystems of Norway and the U.S. (MENU) workshop in Bergen, Norway (2007). Compared energy budget models and trophic network attributes for five high-latitude marine ecosystems, presented results at 2007 ICES and PICES meetings, and collaborated on three publications. 
*	Regular participant in Alaska Fisheries Science Center scientific education and outreach activities, including Pacific Science Center programs, Minorities in Marine Science Undergraduate Program presentations, elementary, secondary, and college educator workshops, internship application review and placement.  


## Professional societies

AFS
AIFRB

## Editorial positions

## Membership on panels

Invited review panelist, Atlantic States Marine Fisheries Commission Atlantic Menhaden stock assessment and ecological reference points, November 2019.

Chaired Baltic Sea multispecies model key run reviews for ICES WGSAM, October 2019.

Invited review panelist, South African Assessment review, November 2018. MSE-based management procedures for 4 species were reviewed over the course of a week by an international panel. (Also invited for 2019 panel, but declined due to conflicts).



## Advisory services

All the Council stuff

ICES WKs

## Workshops organized/chaired; sessions organized/chaired at professional meetings

ICES/PICES 2022 Small Pelagics symposium co-convener

WFC 2021 Climate and Ecosystem section of Horn of Plenty symposium co-convener

2019 National Ecosystem Modeling Workshop (NEMoW) steering committee member; unable to attend due to conflict
with MAFMC EAFM Conceptual Model presentation in December
– Developed NEFSC overview presentation (presented by Scott Large)
– Developed atlantisom tool demonstration (presented by Christine Stawitz)

AFS 2018 ecosystem emergent properties symposium co-convener Atlantic City

AFS 2017 MSE symposium co-convener Tampa

*	Symposium Co-organizer: “Multispecies models (including humans!)—where have we been and where do we need to go?” 2015 American Fisheries Society annual meeting. 

“What Influences Fisheries Production? Comparing the Effects of Environmental, Fishing, and Food Web Forcing Across Large Marine Ecosystems” 2011 American Fisheries Society annual meeting. 

"Bridging the gap between fisheries and marine conservation to advance ecosystem-based management", Society for Conservation Biology's International Marine Conservation Conference (IMCC) 2009. 

"Ecosystem modeling applications in fishery management" 2007 American Fisheries Society annual meeting.

## Teaching and mentoring

**Adjunct professor**  *University of Massachusetts, Dartmouth* 

One current College of William and Mary (Virginia Institute of Marine Science) PhD student committee  
One current UNH PhD student committee

2021  (Exam) PhD student committee at the University of Massachusetts, Dartmouth  
2020  External examiner for a PhD student at University of Iceland  
2019  (Exam) PhD student committee at the University of Massachusetts, Dartmouth   
2019  External examiner for a PhD student at Dalhousie University  
2016	External examiner for a PhD student at the University of Cape Town   
2016	(Exam) PhD student committee at the University of Washington   
2015	(Exam) PhD student committee at Oregon State University  

WPI REU program
Invited speaker for college courses


# Agency Service

## Committees

National MSE WG Chair

MAFMC SSC (here or advisory?)

## Other

# Other relevant information

March 2020 - present: considerable time was needed for dependent care (elementary age kids and elder care). I often used the excused absence for dependent care.

# References


